---
title: Genetic differentiation on flowering time in Cerastium fontanum using a greenhouse
  experiment
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 4
subtitle: Data preparation and first analyses
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(ggeffects)
library(glmmTMB)
library(car)
library(sjPlot)
library(jtools)
library(broom)
library(broom.mixed)
library(knitr)
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
myPalette <- colorRampPalette(brewer.pal(11, "YlOrRd"))
```

# Data preparation

## Read data from Excel files

```{r}
data_exp<-read_excel("data/edited/Cerastium_greenhouse_spring_2022.xlsx", 
                       sheet="extracted_data")
data_exp$treat<-as.factor(data_exp$treat)
data_exp$mother<-as.factor(data_exp$mother)
data_exp$father<-as.factor(data_exp$father)
data_exp$total_n_fl<-as.integer(data_exp$total_n_fl)
data_exp$date50<-as.numeric(data_exp$date50)
```

```{r}
data_parents<-read_excel("data/edited/Cerastium_greenhouse_spring_2022.xlsx", 
                       sheet="extracted_data_parents")
data_parents$treat<-as.factor(data_parents$treat)
```

```{r}
head(data_exp)
head(data_parents)
```

## How many different mothers and fathers?

```{r}
length(levels(data_exp$mother)) # 25 different mothers (shouldn't it be 24?)
length(levels(data_exp$father)) # 24 different mothers
```

## How many unique combinations of mothers and fathers?

```{r}
data_exp<-data_exp%>%mutate(mother_father=paste(mother,father,sep="_"))
```

```{r}
length(unique(data_exp$mother_father)) # 100 (shouldn't it be 96?)
# It is really 99, because one is "NA_NA"
```

## Number of plants from each crossing

```{r}
data_exp%>%group_by(mother_father)%>%summarise(num=n())%>%
  ggplot(aes(x=num))+geom_histogram(bins=12)
```

12 plants from most crossings, but also one with 13 and some with fewer.

# Distributions

```{r}
hist(data_exp$FFD)
hist(data_exp$LFD)
hist(data_exp$date50)
```

All looking quite normal.

# Analyses

## With offspring from crosses

```{r}
model_offpsring_FFD<-glmmTMB(FFD~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father),data_exp)
model_offpsring_LFD<-glmmTMB(LFD~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father),data_exp)
model_offpsring_date50<-glmmTMB(date50~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father),data_exp)
```

```{r eval=FALSE, include=FALSE}
kable(tidy(model_offpsring_FFD),digits=c(3,3,2,3))
kable(tidy(model_offpsring_LFD),digits=c(3,3,2,3))
kable(tidy(model_offpsring_date50),digits=c(3,3,2,3))
```

Save models as HTML table

```{r}
tab_model(model_offpsring_FFD,model_offpsring_LFD,model_offpsring_date50,
          transform=NULL,show.ci=F,show.se=T,show.stat=T,digits=3,
          dv.labels=c("FFD","LFD","date50"),
          file="output/tables/Table_models_offspring.html",
          title="Models offspring")
```

Only effects of treatment are significant for FFD and LFD, also temp_father for date 50.

Plot predictions of significant effects

```{r}
plot(ggpredict(model_offpsring_FFD,terms=c("treat")))
plot(ggpredict(model_offpsring_LFD,terms=c("treat")))
plot(ggpredict(model_offpsring_date50,terms=c("treat")))
plot(ggpredict(model_offpsring_date50,terms=c("temp_father")))
```

Plot predictions of interactions (NS)

```{r}
plot(ggpredict(model_offpsring_FFD,terms=c("temp_mother","treat")))
plot(ggpredict(model_offpsring_LFD,terms=c("temp_mother","treat")))
plot(ggpredict(model_offpsring_date50,terms=c("temp_mother","treat")))
plot(ggpredict(model_offpsring_FFD,terms=c("temp_father","treat")))
plot(ggpredict(model_offpsring_LFD,terms=c("temp_father","treat")))
plot(ggpredict(model_offpsring_date50,terms=c("temp_father","treat")))
```

## With parents

```{r}
model_parents_FFD<-lm(FFD~temp*treat,data_parents)
model_parents_LFD<-lm(LFD~temp*treat,data_parents)
model_parents_date50<-lm(date50~temp*treat,data_parents)
```

```{r eval=FALSE, include=FALSE}
kable(tidy(model_parents_FFD),digits=c(3,3,2,3))
kable(tidy(model_parents_LFD),digits=c(3,3,2,3))
kable(tidy(model_parents_date50),digits=c(3,3,2,3))
```

Save models as HTML table

```{r}
tab_model(model_parents_FFD,model_parents_LFD,model_parents_date50,
          transform=NULL,show.ci=F,show.se=T,show.stat=T,digits=3,
          dv.labels=c("FFD","LFD","date50"),
          file="output/tables/Table_models_parents.html",
          title="Models parents")
```

Only effects of treatment are significant for FFD and LFD, also temp for date 50 (p=0.066 for temp in the model for LFD).

Plot predictions of significant effects

```{r}
plot(ggpredict(model_parents_FFD,terms=c("treat")))
plot(ggpredict(model_parents_LFD,terms=c("treat")))
plot(ggpredict(model_parents_date50,terms=c("treat")))
plot(ggpredict(model_parents_date50,terms=c("temp")))
```

Plot predictions of interactions (NS)

```{r}
plot(ggpredict(model_parents_FFD,terms=c("temp","treat")))
plot(ggpredict(model_parents_LFD,terms=c("temp","treat")))
plot(ggpredict(model_parents_date50,terms=c("temp","treat")))
```

# Session info

```{r}
sessionInfo()
```
