---
title: Genetic differentiation on flowering time in Cerastium fontanum using a greenhouse
  experiment
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 4
  word_document:
    toc: yes
    toc_depth: '4'
subtitle: Data preparation and first analyses
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(readxl)
library(RColorBrewer)
library(ggeffects)
library(glmmTMB)
library(car)
library(sjPlot)
library(jtools)
library(broom)
library(broom.mixed)
library(knitr)
library(lme4)
library(lmerTest)
library(gridExtra)
library(ggthemes)
library(kableExtra)
library(ggpubr)
library(brms)
library(nadiv)
library(bayesplot)
library(matrixStats)
library(tidybayes)
library(cowplot)
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
myPalette <- colorRampPalette(brewer.pal(11, "YlOrRd"))
```

# Data preparation

## Read data from Excel files

```{r}
data_exp<-read_excel("data/edited/Cerastium_greenhouse_spring_2022_editedAV.xlsx", 
                       sheet="extracted_data")
data_exp$treat<-as.factor(data_exp$treat)
data_exp$mother<-as.factor(data_exp$mother)
data_exp$father<-as.factor(data_exp$father)
data_exp$total_n_fl<-as.integer(data_exp$total_n_fl)
data_exp$date50<-as.numeric(data_exp$date50)
data_exp$unique_id<-as.factor(data_exp$unique_id)
data_exp<-data_exp%>%filter(unique_id!="122/144_41") 
# Remove one case where temperature of the parents is not known
```

```{r}
data_selfed<-read_excel("data/edited/Cerastium_greenhouse_spring_2022_editedAV.xlsx", 
                       sheet="extracted_data_parents")
data_selfed$treat<-as.factor(data_selfed$treat)
data_selfed$id_parent<-as.factor(data_selfed$id_parent)
data_selfed<-data_selfed%>%
  mutate(unique_id=paste("selfed",row_number(),sep="_"))
```

```{r}
data_shoots<-read_excel("data/edited/Cerastium_greenhouse_spring_2022_editedAV.xlsx", 
                       sheet="extracted_data_shoots")
data_shoots$unique_id<-as.factor(data_shoots$unique_id)
# Dates say 2022 and that is not true, but should be OK if we need to
# calculate differences
```

```{r}
blocks_mothers<-read_excel("data/edited/crossing_design.xlsx", 
                       sheet="blocks_mothers")
blocks_fathers<-read_excel("data/edited/crossing_design.xlsx", 
                       sheet="blocks_fathers")
```


```{r}
head(data_exp)
head(data_selfed)
head(data_shoots)
head(blocks_mothers)
head(blocks_fathers)
```

## Unique id is unique in all cases --> Fixed

```{r}
data_exp%>%group_by(unique_id)%>%summarise(count=n())%>%filter(count>1)
data_shoots%>%group_by(unique_id)%>%summarise(count=n())%>%filter(count>1)
```

## How many different mothers and fathers?

```{r}
length(levels(data_exp$mother)) # 25 different mothers (shouldn't it be 24?)
length(levels(data_exp$father)) # 24 different mothers
```

## How many unique combinations of mothers and fathers?

```{r}
data_exp<-data_exp%>%mutate(mother_father=paste(mother,father,sep="_"))
```

```{r}
length(unique(data_exp$mother_father)) # 99 (shouldn't it be 96?)
```

## Number of plants from each crossing

```{r}
data_exp%>%group_by(mother_father)%>%summarise(num=n())%>%
  ggplot(aes(x=num))+geom_histogram(bins=12)
```

12 plants from most crossings, but also one with 13 and some with fewer.

## Adding shoot info

```{r}
data_shoots<-data_shoots%>%
  mutate(day_diff=date2-date1,
         mean_sh_h_date1=(sh1_date1+sh2_date1+sh3_date1)/3,
         mean_sh_h_date2=(sh1_date2+sh2_date2+sh3_date2)/3,
         mean_sh_growth=mean_sh_h_date2-mean_sh_h_date1)
```

```{r}
data_exp<-data_exp%>%left_join(data_shoots%>%select(unique_id,mean_sh_growth))
```

## Adding block info to crossings

```{r}
data_exp_blocks<-data_exp%>%
  left_join(blocks_mothers%>%mutate(mother=factor(mother)))%>%
  left_join(blocks_fathers%>%mutate(father=factor(father)))%>%
  # Removing 2 cases where block_mother different from block_father
  filter(block_mother==block_father)%>%
  mutate(block=factor(block_mother))%>%
  select(!(block_mother:block_father))
```

## Add block info to selfed

```{r}
data_selfed_blocks<-data_selfed%>%
  mutate(mother=id_parent,father=id_parent)%>%
  select(-id_parent)%>%
  left_join(blocks_mothers%>%mutate(mother=factor(mother)))%>%
  left_join(blocks_fathers%>%mutate(father=factor(father)))%>%
  mutate(block=factor(ifelse(is.na(block_mother),block_father,block_mother)))%>%
  select(!(block_mother:block_father))
```

# Distributions

```{r}
hist(data_exp$FFD)
hist(data_exp$LFD)
hist(data_exp$date50)
data_exp$duration<-data_exp$LFD-data_exp$FFD
hist(data_exp$duration)
hist(data_exp$mean_sh_growth)
```

All looking quite normal.

# Analyses

## LMs / LMMs with selfed progeny

```{r}
length(levels(data_selfed$id_parent))
```

Selfed offpsring from 42 maternal families. id_parent is mother and father. Temp is temperature of the "grandmother" of the plant.

```{r}
data_selfed%>%group_by(id_parent,treat)%>%
  summarise(mean=mean(FFD),sd=sd(FFD))%>%
  ggplot(aes(x=treat,y=mean,group=id_parent,color=id_parent,label=id_parent))+
  geom_line()+geom_label()+ylab("FFD")+my_theme()
data_selfed%>%group_by(id_parent,treat)%>%
  summarise(mean=mean(LFD),sd=sd(LFD))%>%
  ggplot(aes(x=treat,y=mean,group=id_parent,color=id_parent,label=id_parent))+
  geom_line()+geom_label()+ylab("LFD")+my_theme()
data_selfed%>%group_by(id_parent,treat)%>%
  summarise(mean=mean(date50),sd=sd(date50))%>%
  ggplot(aes(x=treat,y=mean,group=id_parent,color=id_parent,label=id_parent))+
  geom_line()+geom_label()+ylab("date50")+my_theme()
```

Differences in RNs between the 42 maternal lines?

2-3 look quite different in the plots above.

```{r}
model_selfed_FFD_1<-lm(FFD~id_parent*treat,data_selfed)
model_selfed_LFD_1<-lm(LFD~id_parent*treat,data_selfed)
model_selfed_date50_1<-lm(date50~id_parent*treat,data_selfed)
Anova(model_selfed_FFD_1)
Anova(model_selfed_LFD_1)
Anova(model_selfed_date50_1)
```

This suggests that there are no significant differences among individuals in the slope of the RNs. But there are differences in RN elevation.

Differences in RNs depend on origin temperature?

```{r}
model_selfed_FFD_2<-glmmTMB(FFD~temp*treat+(1|id_parent),data_selfed)
model_selfed_LFD_2<-glmmTMB(LFD~temp*treat+(1|id_parent),data_selfed)
model_selfed_date50_2<-glmmTMB(date50~temp*treat+(1|id_parent),data_selfed)
Anova(model_selfed_FFD_2)
Anova(model_selfed_LFD_2)
Anova(model_selfed_date50_2)
```

For LFD and date 50, differences in elevation related to origin temperature. No differences in slope.

```{r}
plot(ggpredict(model_selfed_FFD_2,terms=c("treat")))
plot(ggpredict(model_selfed_LFD_2,terms=c("treat")))
plot(ggpredict(model_selfed_date50_2,terms=c("treat")))
plot(ggpredict(model_selfed_FFD_2,terms=c("temp")))
plot(ggpredict(model_selfed_LFD_2,terms=c("temp")))
plot(ggpredict(model_selfed_date50_2,terms=c("temp")))
```

Individuals with parents from warmer soils have later LFD and date50 at a given temperature than plants with parents from colder soils, i.e. environmental and genetic effects on flowering time are in opposite directions.

## LMMs With progeny from crossings

### Flowering phenology

Including mother, father and block as random factors.

```{r}
model_offpsring_FFD<-glmmTMB(FFD~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father)+(1|block),data_exp_blocks)
model_offpsring_LFD<-glmmTMB(LFD~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father)+(1|block),data_exp_blocks)
model_offpsring_date50<-glmmTMB(date50~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father)+(1|block),data_exp_blocks)
```

Save models as HTML table

```{r}
tab_model(model_offpsring_FFD,model_offpsring_LFD,model_offpsring_date50,
          transform=NULL,show.ci=F,show.se=T,show.stat=T,digits=3,
          dv.labels=c("FFD","LFD","date50"),
          file="output/tables/Table_models_fl_phen.html",
          title="Models flowering phenology")
```

Only effects of treatment are significant for FFD and LFD, also temp_father for date50.

Variance for block is zero, maybe we can remove it.

Plot predictions of significant effects

```{r}
plot(ggpredict(model_offpsring_FFD,terms=c("treat")))
plot(ggpredict(model_offpsring_LFD,terms=c("treat")))
plot(ggpredict(model_offpsring_date50,terms=c("treat")))
plot(ggpredict(model_offpsring_date50,terms=c("temp_father")))
```

Individuals with fathers from warmer soils have later date50 at a given temperature than plants with parents from colder soils, i.e. environmental and genetic effects on flowering time are in opposite directions.

Plot predictions of interactions (NS)

```{r}
plot(ggpredict(model_offpsring_FFD,terms=c("temp_mother","treat")))
plot(ggpredict(model_offpsring_LFD,terms=c("temp_mother","treat")))
plot(ggpredict(model_offpsring_date50,terms=c("temp_mother","treat")))
plot(ggpredict(model_offpsring_FFD,terms=c("temp_father","treat")))
plot(ggpredict(model_offpsring_LFD,terms=c("temp_father","treat")))
plot(ggpredict(model_offpsring_date50,terms=c("temp_father","treat")))
```

#### Get variance components and calculate heritability

Adapted from Elsa's code! Now done only for FFD.

Refit models with lmer (easier to get variance components, might use these overall in the end).
Removed block because of singular fit warnings

```{r}
model_offpsring_FFD_lmer<-lmer(FFD~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father),data_exp_blocks)
model_offpsring_LFD_lmer<-lmer(LFD~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father),data_exp_blocks)
model_offpsring_date50_lmer<-lmer(date50~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father),data_exp_blocks)
```

```{r}
# extract variance components
Variance_FFD <- as.data.frame(VarCorr(model_offpsring_FFD_lmer))[,c(1,4)] 
# Intra-class correlation
PropVar_FFD <- Variance_FFD_treat%>%mutate(propvar=vcov/sum(vcov))
# Proportional variance
PropVar_FFD
```

Heritability and maternal effects

```{r}
# h^2 (paternal effects)
as.numeric(4*subset(PropVar_FFD,grp=="father")[3])
# Because the additive genetic variance, VA,
# is expected to be four times the among pollen‐donor variance 
# (Falconer & Mackay, 1996; Lynch & Walsh, 1998)
```

```{r}
# maternal effects

# Maternal - paternal
# ----------------------
# summed effects

# Because the pollen‐recipient variance component contains a combination of
# genetic and environmental effects, we subtracted the additive genetic 
# (pollen donor) component from the pollen-recipient variance component
# before dividing the resulting estimate by VP to estimate 
# m2 (m2 = (Vpollen recipient − Vpollen donor)/VP).

(Variance_FFD[1,2]-Variance_FFD[2,2])/
  as.numeric(Variance_FFD%>%summarise(sum(vcov)))
```

### Vegetative phenology

Including mother, father and block as random factors.

```{r}
model_offpsring_veg_phen<-glmmTMB(mean_sh_growth~(temp_mother+temp_father)*treat+
                               (1|mother)+(1|father)+(1|block),data_exp_blocks)
```

Save models as HTML table

```{r}
tab_model(model_offpsring_veg_phen,
          transform=NULL,show.ci=F,show.se=T,show.stat=T,digits=3,
          dv.labels=c("Mean shoot growth"),
          file="output/tables/Table_models_veg_phen.html",
          title="Models vegetative phenology")
```

No significant effects.

Variance for block is zero, maybe we can remove it.

## Bayesian models with mother and father

### Crosses

#### date50

```{r eval=FALSE, include=FALSE}
brms_mf_offpsring_date50_full <- brm(
  date50~ 1 + (temp_mother+temp_father)*treat +
    (1|mother) + (1|father) + (1|block) +
  data = data_exp_blocks,
  family = gaussian(),
  chains = 4, cores = 16, iter = 5000, warmup = 1000, thin = 5,
)
save(brms_mf_offpsring_date50_full,
     file = "output/models/brms_mf_offpsring_date50_full.rda")
```

```{r}
load("output/models/brms_mf_offpsring_date50_full.rda")
plot(brms_mf_offpsring_date50_full)
summary(brms_mf_offpsring_date50_full)
```

## Bayesian animal models

Fitting animal models with brms package

Animal model = a form of a mixed model that can estimate the genetic and environmental components of phenotypic variation using pedigree data. Specifically, we used a linear mixed model (LMM) using Markov chain Monte Carlo (MCMC) in the R package brms. These models are particularly useful for estimating genetic variance in unbalanced pedigrees such as in wild populations or a lack of a fully crossed breeding designs. We can use this model to estimate variance components, narrow-sense heritability (h2) and maternal effects.

### Crosses

Using only offspring from crosses.

Create pedigree:

```{r}
data_ped_cross<-data.frame(unique(select(data_exp,unique_id,mother,father)))
```

To be able to fit an animal model, brms needs the relativeness (relationship) matrix of the pedigree and not its inverse (as in other softwares). This can be estimated using the nadiv package.

Create the relativeness (relationship) matrix of the pedigree:

```{r}
Amat_cross <- as.matrix(makeA(prepPed(data_ped_cross,check=T)))
```

#### Both treatments, all effects

##### FFD

The random effect is added to the model with the term (1 | gr(animal, cov = Amat)) which associates the id animal to the matrix of relativeness. 

we will use the default priors so far.

Specify model (with and without block):

```{r eval=FALSE, include=FALSE}
brms_offpsring_FFD_full <- brm(
  FFD~ 1 + (temp_mother+temp_father)*treat + 
    (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) + 
    (1|mother), # To account for maternal effects
  data = data_exp_blocks,
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 5000, warmup = 1000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_offpsring_FFD_full, 
     file = "output/models/brms_offpsring_FFD_full.rda")
```

Model diagnostics, examine variance estimate and distributions:

```{r}
load("output/models/brms_offpsring_FFD_full.rda")
plot(brms_offpsring_FFD_full)
summary(brms_offpsring_FFD_full)
```

Rhat provides information on the estimate convergence. If it’s greater than 1, the chains have not yet converged and it will be required to run more iterations and/or set stronger priors.

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_offpsring_FFD_full))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_offpsring_FFD_full))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

Look at fixed effects:

```{r}
conditional_effects(brms_offpsring_FFD_full)
```

The posterior distribution of the treat term does not overlap zero. Thus, we can infer that treatment has an effect on FFD.

Calculate heritability and maternal effects:

Not including fixed effects in phenotypic variance

```{r}
v_animal_offpsring_FFD_full<-(VarCorr(brms_offpsring_FFD_full,
                                        summary=FALSE)$unique_id$sd)^2
v_block_offpsring_FFD_full<-(VarCorr(brms_offpsring_FFD_full,
                                     summary=FALSE)$block$sd)^2
v_mother_offpsring_FFD_full<-(VarCorr(brms_offpsring_FFD_full,
                                      summary=FALSE)$mother$sd)^2
v_r_offpsring_FFD_full<-(VarCorr(brms_offpsring_FFD_full,
                                   summary=FALSE)$residual$sd)^2
h1_offpsring_FFD_full<-as.mcmc(v_animal_offpsring_FFD_full/
                                  (v_animal_offpsring_FFD_full+
                                     v_block_offpsring_FFD_full+
                                     v_mother_offpsring_FFD_full+
                                     v_r_offpsring_FFD_full))
m1_offpsring_FFD_full<-as.mcmc(v_mother_offpsring_FFD_full/
                                  (v_animal_offpsring_FFD_full+
                                     v_block_offpsring_FFD_full+
                                     v_mother_offpsring_FFD_full+
                                     v_r_offpsring_FFD_full))
summary(h1_offpsring_FFD_full)
summary(m1_offpsring_FFD_full)
plot(h1_offpsring_FFD_full)
plot(m1_offpsring_FFD_full)
```

We could compute the variance arising from the fixed effects as the variance of the predicted values (Nakagawa & Schielzeth 2013; de Villemereuil et al. 2018), and add it to the denominator of heritability. Then the total phenotypic variance would consist of (1) Va, (2) Vf, the variance from fixed effects, (3) Vr, the variance of additional random effects, and (4) Vre, the residual variance component.

However, de Villemereuil et al. (2018) suggests to include only ‘natural’ fixed effects (e.g. sex, and not treatment). So I don't think we should include the variance of the fixed effect of treatment in the denominator here. But we could include the variance related to origin temperature (as this is a "natural" fixed effect).

Including fixed effects (temperatures of origin but not treatment, according to de Villemereuil et al. 2018) in phenotypic variance

```{r}
v_fixed_offpsring_FFD_full<-rowVars(
  posterior_linpred(brms_offpsring_FFD_full,
                    newdata=data_exp_blocks%>%
                      mutate(treat="unheated"),re_formula = NA))
# Variance of fixed effects (temp only)
# Modified newdata because it is the variance of fixed effect
# assuming individuals only differ in temps_mother and temp_father
# Idea from code of Dahirel et al. 2021
# https://github.com/mdahirel/cepaea-personality-2017/blob/main/R/cepaea-2017-script.Rmd
v_animal_offpsring_FFD_full<-(VarCorr(brms_offpsring_FFD_full,
                                        summary=FALSE)$unique_id$sd)^2
v_block_offpsring_FFD_full<-(VarCorr(brms_offpsring_FFD_full,
                                     summary=FALSE)$block$sd)^2
v_mother_offpsring_FFD_full<-(VarCorr(brms_offpsring_FFD_full,
                                      summary=FALSE)$mother$sd)^2
v_r_offpsring_FFD_full<-(VarCorr(brms_offpsring_FFD_full,
                                   summary=FALSE)$residual$sd)^2
h2_offpsring_FFD_full<-as.mcmc(v_animal_offpsring_FFD_full/
                                  (v_fixed_offpsring_FFD_full+
                                     v_animal_offpsring_FFD_full+
                                     v_block_offpsring_FFD_full+
                                     v_mother_offpsring_FFD_full+
                                     v_r_offpsring_FFD_full))
m2_offpsring_FFD_full<-as.mcmc(v_mother_offpsring_FFD_full/
                                  (v_fixed_offpsring_FFD_full+
                                     v_animal_offpsring_FFD_full+
                                     v_block_offpsring_FFD_full+
                                     v_mother_offpsring_FFD_full+
                                     v_r_offpsring_FFD_full))
prop_var_temp_offpsring_FFD_full<-as.mcmc(v_fixed_offpsring_FFD_full/
                                  (v_fixed_offpsring_FFD_full+
                                     v_animal_offpsring_FFD_full+
                                     v_block_offpsring_FFD_full+
                                     v_mother_offpsring_FFD_full+
                                     v_r_offpsring_FFD_full))
summary(h2_offpsring_FFD_full)
summary(m2_offpsring_FFD_full)
summary(prop_var_temp_offpsring_FFD_full)
plot(h2_offpsring_FFD_full)
plot(m2_offpsring_FFD_full)
plot(prop_var_temp_offpsring_FFD_full)
```

Compare the two values of heritability and maternal effects:

```{r}
mean(h1_offpsring_FFD_full)
mean(h2_offpsring_FFD_full)
mean(m1_offpsring_FFD_full)
mean(m2_offpsring_FFD_full)
```

Heritabilities and maternal effects are only slightly lower when including fixed effects (= temperatures of origin contribute little to phenotypic variance. I guess this makes sense as there were no significant effects of origin temperatures).

###### Variance components

Ideas from code of Dahirel et al. 2021
https://github.com/mdahirel/cepaea-personality-2017/blob/main/R/cepaea-2017-script.Rmd

```{r}
varcomps_offpsring_FFD_full <- tibble(
  beta0 = posterior_samples(brms_offpsring_FFD_full,pars="b_Intercept")[,1],
  VF= rowVars(posterior_linpred(brms_offpsring_FFD_full, re_formula = NA)),
  VFtemp= rowVars(posterior_linpred(brms_offpsring_FFD_full, 
                                     newdata=data_exp_blocks%>%
                                       mutate(treat="unheated"), 
                                     re_formula = NA)),
  VI = (VarCorr(brms_offpsring_FFD_full,summary=F)$unique_id$sd[,"Intercept"])^2,
  VM = (VarCorr(brms_offpsring_FFD_full,summary=F)$mother$sd[,"Intercept"])^2,
  VB= VarCorr(brms_offpsring_FFD_full,summary=FALSE)$block$sd[,"Intercept"]^2,
  VR=(posterior_samples(brms_offpsring_FFD_full,pars="sigma")[,1])^2
  ) %>% 
  mutate(VP = select(.,c(VF,VI,VM,VB,VR)) %>% rowSums()) %>% 
  mutate(.iteration = 1) %>% mutate(.iteration = cumsum(.iteration))
```

```{r}
varcomps_offpsring_FFD_full %>% 
  pivot_longer(cols=c(VF:VR),names_to = "varcomp" , values_to = "variance") %>% 
  mutate(varcomp=fct_relevel(varcomp,c("VR","VI","VM","VB","VFtemp","VF"))) %>% 
  ggplot()+
  stat_halfeye(aes(y=varcomp,x=100*variance/VP),
                point_interval = mean_hdi,.width = 0.001,
                normalize="xy")+
  scale_x_continuous(name="Variance explained (%)",
                     lim=c(0,100),breaks=seq(from=0,to=100,by=10))+
  scale_y_discrete(name="Variance component",
                   labels=c(expression(italic(V[Residual])),
                            expression(italic(V[Individual])),
                            expression(italic(V[Mother])),
                            expression(italic(V[Block])),
                            expression(italic(V[Fixed(temp)])),
                            expression(italic(V[Fixed]))
                            ))+
  my_theme()+
  background_grid(colour.major = "grey95",colour.minor = "grey95")+
  ggtitle("Partitioning of total phenotypic variation for FFD")+
  labs(caption = expression(paste("Total phenotypic variance =",italic(V[Fixed]),
                           " + ", italic(V[Block])," + ", italic(V[Mother]),
                           " + ", italic(V[Individual]),
                           " + ", italic(V[Residual])
                           )))
```

###### Extract BLUPs

```{r}
BLUPs_brms_offpsring_FFD_full<-as.data.frame(ranef(
  brms_offpsring_FFD_full)$unique_id)[1]%>%
  rownames_to_column()%>%
  rename(unique_id=rowname,BLUP_intercept=Estimate.Intercept)
```

```{r}
ggplot(BLUPs_brms_offpsring_FFD_full, aes(unique_id, BLUP_intercept)) + 
  geom_point(aes(group = unique_id, colour = unique_id), size = 4, alpha = 0.5) + 
  ylab("BLUP intercept estimate") +
  geom_hline(yintercept = 0, lty = 2) + my_theme() +
  theme(axis.text.x = element_blank())
```

```{r}
data_exp_blocks<-data_exp_blocks%>%
  left_join(BLUPs_brms_offpsring_FFD_full)
```

```{r}
model_BLUPs<-lm(BLUP_intercept~temp_mother*temp_father,data_exp_blocks)
summary(model_BLUPs)
plot(ggpredict(model_BLUPs,terms=c("temp_father","temp_mother")))
```

Why interaction?

Now not sure why I did this analysis with BLUPs anymore!

##### LFD

```{r eval=FALSE, include=FALSE}
brms_offpsring_LFD_full <- brm(
  LFD~ 1 + (temp_mother+temp_father)*treat +
    (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) +
    (1|mother), # To account for maternal effects
  data = data_exp_blocks,
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 5000, warmup = 1000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_offpsring_LFD_full,
     file = "output/models/brms_offpsring_LFD_full.rda")
```

```{r}
load("output/models/brms_offpsring_LFD_full.rda")
plot(brms_offpsring_LFD_full)
summary(brms_offpsring_LFD_full)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_offpsring_LFD_full))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_offpsring_LFD_full))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

```{r}
conditional_effects(brms_offpsring_LFD_full)
```

##### date50

```{r eval=FALSE, include=FALSE}
brms_offpsring_date50_full <- brm(
  date50~ 1 + (temp_mother+temp_father)*treat +
    (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) +
    (1|mother), # To account for maternal effects
  data = data_exp_blocks,
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 5000, warmup = 1000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_offpsring_date50_full,
     file = "output/models/brms_offpsring_date50_full.rda")
```

```{r}
load("output/models/brms_offpsring_date50_full.rda")
plot(brms_offpsring_date50_full)
summary(brms_offpsring_date50_full)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_offpsring_date50_full))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_offpsring_date50_full))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

```{r}
conditional_effects(brms_offpsring_date50_full)
```

###### Variance components

Ideas from code of Dahirel et al. 2021
https://github.com/mdahirel/cepaea-personality-2017/blob/main/R/cepaea-2017-script.Rmd

```{r}
varcomps_offpsring_date50_full <- tibble(
  beta0 = posterior_samples(brms_offpsring_date50_full,pars="b_Intercept")[,1],
  VF= rowVars(posterior_linpred(brms_offpsring_date50_full, re_formula = NA)),
  VFtemp= rowVars(posterior_linpred(brms_offpsring_date50_full, 
                                     newdata=data_exp_blocks%>%
                                       mutate(treat="unheated"), 
                                     re_formula = NA)),
  VI = (VarCorr(brms_offpsring_date50_full,summary=F)$unique_id$sd[,"Intercept"])^2,
  VM = (VarCorr(brms_offpsring_date50_full,summary=F)$mother$sd[,"Intercept"])^2,
  VB= VarCorr(brms_offpsring_date50_full,summary=FALSE)$block$sd[,"Intercept"]^2,
  VR=(posterior_samples(brms_offpsring_date50_full,pars="sigma")[,1])^2
  ) %>% 
  mutate(VP = select(.,c(VF,VI,VM,VB,VR)) %>% rowSums()) %>% 
  mutate(.iteration = 1) %>% mutate(.iteration = cumsum(.iteration))
```

```{r}
varcomps_offpsring_date50_full %>% 
  pivot_longer(cols=c(VF:VR),names_to = "varcomp" , values_to = "variance") %>% 
  mutate(varcomp=fct_relevel(varcomp,c("VR","VI","VM","VB","VFtemp","VF"))) %>% 
  ggplot()+
  stat_halfeye(aes(y=varcomp,x=100*variance/VP),
                point_interval = mean_hdi,.width = 0.001,
                normalize="xy")+
  scale_x_continuous(name="Variance explained (%)",
                     lim=c(0,100),breaks=seq(from=0,to=100,by=10))+
  scale_y_discrete(name="Variance component",
                   labels=c(expression(italic(V[Residual])),
                            expression(italic(V[Individual])),
                            expression(italic(V[Mother])),
                            expression(italic(V[Block])),
                            expression(italic(V[Fixed(temp)])),
                            expression(italic(V[Fixed]))
                            ))+
  my_theme()+
  background_grid(colour.major = "grey95",colour.minor = "grey95")+
  ggtitle("Partitioning of total phenotypic variation for date50")+
  labs(caption = expression(paste("Total phenotypic variance =",italic(V[Fixed]),
                           " + ", italic(V[Block])," + ", italic(V[Mother]),
                           " + ", italic(V[Individual]),
                           " + ", italic(V[Residual])
                           )))
```

#### Both treatments, no origin temperatures

##### FFD

Specify model:

```{r eval=FALSE, include=FALSE}
brms_offpsring_FFD_notemp <- brm(
  FFD~ 1 + treat + 
    (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) + 
    (1|mother), # To account for maternal effects
  data = data_exp_blocks,
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 10000, warmup = 3000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_offpsring_FFD_notemp, 
     file = "output/models/brms_offpsring_FFD_notemp.rda")
```

```{r}
load("output/models/brms_offpsring_FFD_notemp.rda")
plot(brms_offpsring_FFD_notemp)
summary(brms_offpsring_FFD_notemp)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_offpsring_FFD_notemp))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_offpsring_FFD_notemp))+
            my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

Look at fixed effects:

```{r}
conditional_effects(brms_offpsring_FFD_notemp)
```

Calculate heritability and maternal effects:

```{r}
v_animal_offpsring_FFD_notemp<-(VarCorr(brms_offpsring_FFD_notemp,
                                        summary=FALSE)$unique_id$sd)^2
v_block_offpsring_FFD_notemp<-(VarCorr(brms_offpsring_FFD_notemp,
                                     summary=FALSE)$block$sd)^2
v_mother_offpsring_FFD_notemp<-(VarCorr(brms_offpsring_FFD_notemp,
                                      summary=FALSE)$mother$sd)^2
v_r_offpsring_FFD_notemp<-(VarCorr(brms_offpsring_FFD_notemp,
                                   summary=FALSE)$residual$sd)^2
h_offpsring_FFD_notemp<-as.mcmc(v_animal_offpsring_FFD_notemp/
                                  (v_animal_offpsring_FFD_notemp+
                                     v_block_offpsring_FFD_notemp+
                                     v_mother_offpsring_FFD_notemp+
                                     v_r_offpsring_FFD_notemp))
m_offpsring_FFD_notemp<-as.mcmc(v_mother_offpsring_FFD_notemp/
                                  (v_animal_offpsring_FFD_notemp+
                                     v_block_offpsring_FFD_notemp+
                                     v_mother_offpsring_FFD_notemp+
                                     v_r_offpsring_FFD_notemp))
summary(h_offpsring_FFD_notemp)
summary(m_offpsring_FFD_notemp)
plot(h_offpsring_FFD_notemp)
plot(m_offpsring_FFD_notemp)
```

Heritability values are quite similar to the ones from previous model.
Including temperature of origin in the model does not matter a lot for calculation of heritability.

##### LFD (model with problems)

Warning message:
There were 8 divergent transitions after warmup. Increasing adapt_delta above 0.999 may help. See http://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup 

```{r eval=FALSE, include=FALSE}
brms_offpsring_LFD_notemp <- brm(
  LFD~ 1 + treat +
    (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) +
    (1|mother), # To account for maternal effects
  data = data_exp_blocks,
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 25000, warmup = 5000, thin = 5,
  control = list(adapt_delta = 0.999)
)
save(brms_offpsring_LFD_notemp,
     file = "output/models/brms_offpsring_LFD_notemp.rda")
```

```{r}
load("output/models/brms_offpsring_LFD_notemp.rda")
plot(brms_offpsring_LFD_notemp)
summary(brms_offpsring_LFD_notemp)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_offpsring_LFD_notemp))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_offpsring_LFD_notemp))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

```{r}
conditional_effects(brms_offpsring_LFD_notemp)
```


##### date50 (model with problems)

Warning message:
There were 158 divergent transitions after warmup. Increasing adapt_delta above 0.99 may help. See http://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup

Warning messages:
1: Parts of the model have not converged (some Rhats are > 1.05). Be careful when analysing the results! We recommend running more iterations and/or setting stronger priors. 
2: There were 2192 divergent transitions after warmup. Increasing adapt_delta above 0.999 may help. See http://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup 

```{r eval=FALSE, include=FALSE}
brms_offpsring_date50_notemp <- brm(
  date50~ 1 + treat +
    (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) +
    (1|mother), # To account for maternal effects
  data = data_exp_blocks,
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 25000, warmup = 5000, thin = 5,
  control = list(adapt_delta = 0.999)
)
save(brms_offpsring_date50_notemp,
     file = "output/models/brms_offpsring_date50_notemp.rda")
```

```{r}
load("output/models/brms_offpsring_date50_notemp.rda")
plot(brms_offpsring_date50_notemp)
summary(brms_offpsring_date50_notemp)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_offpsring_date50_notemp))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_offpsring_date50_notemp))+
            my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

```{r}
conditional_effects(brms_offpsring_date50_notemp)
```

#### Without treatments, without origin temperatures

##### FFD

Specify model:

```{r eval=FALSE, include=FALSE}
brms_offpsring_FFD_notreat_notemp <- brm(
  FFD~ 1 +  
    (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) + 
    (1|mother), # To account for maternal effects
  data = data_exp_blocks,
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 10000, warmup = 3000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_offpsring_FFD_notreat_notemp, 
     file = "output/models/brms_offpsring_FFD_notreat_notemp.rda")
```

```{r}
load("output/models/brms_offpsring_FFD_notreat_notemp.rda")
plot(brms_offpsring_FFD_notreat_notemp)
summary(brms_offpsring_FFD_notreat_notemp)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_offpsring_FFD_notreat_notemp))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_offpsring_FFD_notreat_notemp))+
            my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

Calculate heritability and maternal effects:

```{r}
v_animal_offpsring_FFD_notreat_notemp<-(VarCorr(brms_offpsring_FFD_notreat_notemp,
                                        summary=FALSE)$unique_id$sd)^2
v_block_offpsring_FFD_notreat_notemp<-(VarCorr(brms_offpsring_FFD_notreat_notemp,
                                     summary=FALSE)$block$sd)^2
v_mother_offpsring_FFD_notreat_notemp<-(VarCorr(brms_offpsring_FFD_notreat_notemp,
                                      summary=FALSE)$mother$sd)^2
v_r_offpsring_FFD_notreat_notemp<-(VarCorr(brms_offpsring_FFD_notreat_notemp,
                                   summary=FALSE)$residual$sd)^2
h_offpsring_FFD_notreat_notemp<-as.mcmc(v_animal_offpsring_FFD_notreat_notemp/
                                  (v_animal_offpsring_FFD_notreat_notemp+
                                     v_block_offpsring_FFD_notreat_notemp+
                                     v_mother_offpsring_FFD_notreat_notemp+
                                     v_r_offpsring_FFD_notreat_notemp))
m_offpsring_FFD_notreat_notemp<-as.mcmc(v_mother_offpsring_FFD_notreat_notemp/
                                  (v_animal_offpsring_FFD_notreat_notemp+
                                     v_block_offpsring_FFD_notreat_notemp+
                                     v_mother_offpsring_FFD_notreat_notemp+
                                     v_r_offpsring_FFD_notreat_notemp))
summary(h_offpsring_FFD_notreat_notemp)
summary(m_offpsring_FFD_notreat_notemp)
plot(h_offpsring_FFD_notreat_notemp)
plot(m_offpsring_FFD_notreat_notemp)
```

Heritability decreased compared to values from previous models. 
This is because including the fixed effect (treatment) decreases the residual variance.
So in this model without fixed effects, the residual variance is higher, and thus heritability is lower.

##### LFD

```{r eval=FALSE, include=FALSE}
brms_offpsring_LFD_notreat_notemp <- brm(
  LFD~ 1 +
    (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) +
    (1|mother), # To account for maternal effects
  data = data_exp_blocks,
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 10000, warmup = 3000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_offpsring_LFD_notreat_notemp,
     file = "output/models/brms_offpsring_LFD_notreat_notemp.rda")
```

```{r}
load("output/models/brms_offpsring_LFD_notreat_notemp.rda")
plot(brms_offpsring_LFD_notreat_notemp)
summary(brms_offpsring_LFD_notreat_notemp)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_offpsring_LFD_notreat_notemp))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_offpsring_LFD_notreat_notemp))+
            my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

##### date50

```{r eval=FALSE, include=FALSE}
brms_offpsring_date50_notreat_notemp <- brm(
  date50~ 1 +
    (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) +
    (1|mother), # To account for maternal effects
  data = data_exp_blocks,
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 15000, warmup = 6000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_offpsring_date50_notreat_notemp,
     file = "output/models/brms_offpsring_date50_notreat_notemp.rda")
```

```{r}
load("output/models/brms_offpsring_date50_notreat_notemp.rda")
plot(brms_offpsring_date50_notreat_notemp)
summary(brms_offpsring_date50_notreat_notemp)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_offpsring_date50_notreat_notemp))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_offpsring_date50_notreat_notemp))+
            my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

### Crosses + selfed

Using offspring form crosses + selfed progeny.

Create pedigree:

```{r}
data_ped_all<-rbind(
  data.frame(unique(select(data_exp,unique_id,mother,father))),
  data.frame(select(data_selfed,unique_id,id_parent)%>%
               mutate(mother=id_parent,father=id_parent)%>%
               select(-id_parent))
)
```

Create the relativeness (relationship) matrix of the pedigree:

```{r}
Amat_all<-as.matrix(makeA(prepPed(data_ped_all,check=T)))
```

Merge data crossings and selfed:

```{r}
data_all<-rbind(
  data_exp_blocks%>%select(unique_id,block,mother,father,
                           temp_mother,temp_father,treat,FFD,LFD,date50)%>%
    # Not including mean_sh_growth so far cause not measured in selfed
    mutate(category=factor("crossed")),
  data_selfed_blocks%>%mutate(temp_mother=temp,temp_father=temp)%>%
    select(unique_id,block,mother,father,
           temp_mother,temp_father,treat,FFD,LFD,date50)%>%
    mutate(category=factor("selfed"))
)
```

#### Both treatments, all effects

##### FFD

The random effect is added to the model with the term (1 | gr(animal, cov = Amat) which associates the id animal to the matrix of relativeness. 

we will use the default priors so far.

Specify model: 

```{r eval=FALSE, include=FALSE}
brms_all_FFD_full <- brm(
  FFD~ 1 + (temp_mother+temp_father)*treat + 
    (1 | gr(unique_id, cov = Amat_all)) + (1 | block) + 
    (1|mother), # To account for maternal effects
  data = data_all,
  data2 = list(Amat_all = Amat_all),
  chains = 4, cores = 16, iter = 5000, warmup = 1000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_all_FFD_full, 
     file = "output/models/brms_all_FFD_full.rda")
```

Model diagnostics, examine variance estimate and distributions:

```{r}
load("output/models/brms_all_FFD_full.rda")
plot(brms_all_FFD_full)
summary(brms_all_FFD_full)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_all_FFD_full))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_all_FFD_full))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

Look at fixed effects:

```{r}
conditional_effects(brms_all_FFD_full)
```

The posterior distribution of the treat term does not overlap zero. Thus, we can infer that treatment has an effect on FFD.

Calculate heritability and maternal effects:

Not including fixed effects in phenotypic variance

```{r}
v_animal_all_FFD_full<-(VarCorr(brms_all_FFD_full,
                                        summary=FALSE)$unique_id$sd)^2
v_block_all_FFD_full<-(VarCorr(brms_all_FFD_full,
                                     summary=FALSE)$block$sd)^2
v_mother_all_FFD_full<-(VarCorr(brms_all_FFD_full,
                                      summary=FALSE)$mother$sd)^2
v_r_all_FFD_full<-(VarCorr(brms_all_FFD_full,
                                   summary=FALSE)$residual$sd)^2
h1_all_FFD_full<-as.mcmc(v_animal_all_FFD_full/
                                  (v_animal_all_FFD_full+
                                     v_block_all_FFD_full+
                                     v_mother_all_FFD_full+
                                     v_r_all_FFD_full))
m1_all_FFD_full<-as.mcmc(v_mother_all_FFD_full/
                                  (v_animal_all_FFD_full+
                                     v_block_all_FFD_full+
                                     v_mother_all_FFD_full+
                                     v_r_all_FFD_full))
summary(h1_all_FFD_full)
summary(m1_all_FFD_full)
plot(h1_all_FFD_full)
plot(m1_all_FFD_full)
```

Including fixed effects (temperatures of origin but not treatment, according to de Villemereuil et al. 2018) in phenotypic variance

```{r}
v_fixed_all_FFD_full<-rowVars(
  posterior_linpred(brms_all_FFD_full,
                    newdata=data_exp_blocks%>%
                      mutate(treat="unheated"),re_formula = NA))
# Variance of fixed effects (temp only)
# Modified newdata because it is the variance of fixed effect
# assuming individuals only differ in temps_mother and temp_father
# Idea from code of Dahirel et al. 2021
# https://github.com/mdahirel/cepaea-personality-2017/blob/main/R/cepaea-2017-script.Rmd
v_animal_all_FFD_full<-(VarCorr(brms_all_FFD_full,
                                        summary=FALSE)$unique_id$sd)^2
v_block_all_FFD_full<-(VarCorr(brms_all_FFD_full,
                                     summary=FALSE)$block$sd)^2
v_mother_all_FFD_full<-(VarCorr(brms_all_FFD_full,
                                      summary=FALSE)$mother$sd)^2
v_r_all_FFD_full<-(VarCorr(brms_all_FFD_full,
                                   summary=FALSE)$residual$sd)^2
h2_all_FFD_full<-as.mcmc(v_animal_all_FFD_full/
                                  (v_fixed_all_FFD_full+
                                     v_animal_all_FFD_full+
                                     v_block_all_FFD_full+
                                     v_mother_all_FFD_full+
                                     v_r_all_FFD_full))
m2_all_FFD_full<-as.mcmc(v_mother_all_FFD_full/
                                  (v_fixed_all_FFD_full+
                                     v_animal_all_FFD_full+
                                     v_block_all_FFD_full+
                                     v_mother_all_FFD_full+
                                     v_r_all_FFD_full))
summary(h2_all_FFD_full)
summary(m2_all_FFD_full)
plot(h2_all_FFD_full)
plot(m2_all_FFD_full)
```

##### LFD (model with problems)

Warning messages:
1: Parts of the model have not converged (some Rhats are > 1.05). Be careful when analysing the results! We recommend running more iterations and/or setting stronger priors. 
2: There were 152 divergent transitions after warmup. Increasing adapt_delta above 0.8 may help. See http://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup 

```{r eval=FALSE, include=FALSE}
brms_all_LFD_full <- brm(
  LFD~ 1 + (temp_mother+temp_father)*treat + 
    (1 | gr(unique_id, cov = Amat_all)) + (1 | block) + 
    (1|mother), # To account for maternal effects
  data = data_all,
  data2 = list(Amat_all = Amat_all),
  chains = 4, cores = 16, iter = 5000, warmup = 1000, thin = 5
)
save(brms_all_LFD_full, 
     file = "output/models/brms_all_LFD_full.rda")
```

Model diagnostics, examine variance estimate and distributions:

```{r}
load("output/models/brms_all_LFD_full.rda")
plot(brms_all_LFD_full)
summary(brms_all_LFD_full)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_all_LFD_full))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_all_LFD_full))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

##### date50 (model with problems)

Warning messages:
1: Parts of the model have not converged (some Rhats are > 1.05). Be careful when analysing the results! We recommend running more iterations and/or setting stronger priors. 
2: There were 165 divergent transitions after warmup. Increasing adapt_delta above 0.99 may help. See http://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup 

```{r eval=FALSE, include=FALSE}
brms_all_date50_full <- brm(
  date50~ 1 + (temp_mother+temp_father)*treat + 
    (1 | gr(unique_id, cov = Amat_all)) + (1 | block) + 
    (1|mother), # To account for maternal effects
  data = data_all,
  data2 = list(Amat_all = Amat_all),
  chains = 4, cores = 16, iter = 10000, warmup = 3000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_all_date50_full, 
     file = "output/models/brms_all_date50_full.rda")
```

Model diagnostics, examine variance estimate and distributions:

```{r}
load("output/models/brms_all_date50_full.rda")
plot(brms_all_date50_full)
summary(brms_all_date50_full)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_all_date50_full))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_all_date50_full))+my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

#### Both treatments, no origin temperatures

##### FFD

Specify model:

```{r eval=FALSE, include=FALSE}
brms_all_FFD_notemp <- brm(
  FFD~ 1 + treat + 
    (1 | gr(unique_id, cov = Amat_all)) + (1 | block) + 
    (1|mother), # To account for maternal effects
  data = data_all,
  data2 = list(Amat_all = Amat_all),
  family = gaussian(),
  chains = 4, cores = 16, iter = 10000, warmup = 3000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_all_FFD_notemp, 
     file = "output/models/brms_all_FFD_notemp.rda")
```

```{r}
load("output/models/brms_all_FFD_notemp.rda")
plot(brms_all_FFD_notemp)
summary(brms_all_FFD_notemp)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_all_FFD_notemp))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_all_FFD_notemp))+
            my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

Look at fixed effects:

```{r}
conditional_effects(brms_all_FFD_notemp)
```

Calculate heritability and maternal effects:

```{r}
v_animal_all_FFD_notemp<-(VarCorr(brms_all_FFD_notemp,
                                        summary=FALSE)$unique_id$sd)^2
v_block_all_FFD_notemp<-(VarCorr(brms_all_FFD_notemp,
                                     summary=FALSE)$block$sd)^2
v_mother_all_FFD_notemp<-(VarCorr(brms_all_FFD_notemp,
                                      summary=FALSE)$mother$sd)^2
v_r_all_FFD_notemp<-(VarCorr(brms_all_FFD_notemp,
                                   summary=FALSE)$residual$sd)^2
h_all_FFD_notemp<-as.mcmc(v_animal_all_FFD_notemp/
                                  (v_animal_all_FFD_notemp+
                                     v_block_all_FFD_notemp+
                                     v_mother_all_FFD_notemp+
                                     v_r_all_FFD_notemp))
m_all_FFD_notemp<-as.mcmc(v_mother_all_FFD_notemp/
                                  (v_animal_all_FFD_notemp+
                                     v_block_all_FFD_notemp+
                                     v_mother_all_FFD_notemp+
                                     v_r_all_FFD_notemp))
summary(h_all_FFD_notemp)
summary(m_all_FFD_notemp)
plot(h_all_FFD_notemp)
plot(m_all_FFD_notemp)
```

##### RUN: LFD

##### RUN: date50

#### Without treatments, without origin temperatures

##### FFD

Specify model:

```{r eval=FALSE, include=FALSE}
brms_all_FFD_notreat_notemp <- brm(
  FFD~ 1 +  
    (1 | gr(unique_id, cov = Amat_all)) + (1 | block) + 
    (1|mother), # To account for maternal effects
  data = data_all,
  data2 = list(Amat_all = Amat_all),
  family = gaussian(),
  chains = 4, cores = 16, iter = 10000, warmup = 3000, thin = 5,
  control = list(adapt_delta = 0.99)
)
save(brms_all_FFD_notreat_notemp, 
     file = "output/models/brms_all_FFD_notreat_notemp.rda")
```

```{r}
load("output/models/brms_all_FFD_notreat_notemp.rda")
plot(brms_all_FFD_notreat_notemp)
summary(brms_all_FFD_notreat_notemp)
```

```{r}
cowplot::plot_grid(mcmc_rhat_hist(brms::rhat(brms_all_FFD_notreat_notemp))+
                     my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          mcmc_neff_hist(neff_ratio(brms_all_FFD_notreat_notemp))+
            my_theme_legend()+
            theme(legend.position="top")+
            theme(legend.text=element_text(size=10)),
          labels = c("A)","B)"),label_fontfamily="serif")
```

Calculate heritability and maternal effects:

```{r}
v_animal_all_FFD_notreat_notemp<-(VarCorr(brms_all_FFD_notreat_notemp,
                                        summary=FALSE)$unique_id$sd)^2
v_block_all_FFD_notreat_notemp<-(VarCorr(brms_all_FFD_notreat_notemp,
                                     summary=FALSE)$block$sd)^2
v_mother_all_FFD_notreat_notemp<-(VarCorr(brms_all_FFD_notreat_notemp,
                                      summary=FALSE)$mother$sd)^2
v_r_all_FFD_notreat_notemp<-(VarCorr(brms_all_FFD_notreat_notemp,
                                   summary=FALSE)$residual$sd)^2
h_all_FFD_notreat_notemp<-as.mcmc(v_animal_all_FFD_notreat_notemp/
                                  (v_animal_all_FFD_notreat_notemp+
                                     v_block_all_FFD_notreat_notemp+
                                     v_mother_all_FFD_notreat_notemp+
                                     v_r_all_FFD_notreat_notemp))
m_all_FFD_notreat_notemp<-as.mcmc(v_mother_all_FFD_notreat_notemp/
                                  (v_animal_all_FFD_notreat_notemp+
                                     v_block_all_FFD_notreat_notemp+
                                     v_mother_all_FFD_notreat_notemp+
                                     v_r_all_FFD_notreat_notemp))
summary(h_all_FFD_notreat_notemp)
summary(m_all_FFD_notreat_notemp)
plot(h_all_FFD_notreat_notemp)
plot(m_all_FFD_notreat_notemp)
```

##### RUN: LFD

##### RUN: date50


### BIVARIATE - not working!

```{r}
data_wider<-data_exp_blocks%>%
  select(mother,father,temp_mother,temp_father,treat,FFD,unique_id,block)%>%
  pivot_wider(names_from="treat",values_from="FFD")
```


```{r eval=FALSE, include=FALSE}
bf_unheated <- bf(unheated ~ 1 + (1 | a | gr(unique_id, cov = Amat_cross)))
bf_heatmat <- bf(heatmat ~ 1 + (1 | a | gr(unique_id, cov = Amat_cross)))

brms_biv_offspring_FFD <- brm(
  bf_unheated + bf_heatmat + set_rescor(TRUE),
  data = data_wider,
  data2 = list(Amat_cross = Amat_cross),
  chains = 2, cores = 2, iter = 1000
)
save(brms_biv_offspring_FFD, file = "output/models/brms_biv_offspring_FFD.rda")

# brms_offpsring_FFD_full <- brm(
#   FFD~ 1 + (temp_mother+temp_father)*treat + 
#     (1 | gr(unique_id, cov = Amat_cross)) + (1 | block) + 
#     (1|mother), # To account for maternal effects
#   data = data_exp_blocks,
#   data2 = list(Amat_cross = Amat_cross),
#   family = gaussian(),
#   chains = 4, cores = 16, iter = 5000, warmup = 1000, thin = 5,
#   control = list(adapt_delta = 0.99)
# )

```

```{r eval=FALSE, include=FALSE}
brms_biv_offspring_FFD <- brm(
  bf(mvbind(unheated, heatmat) ~ (1|p|gr(unique_id, cov = Amat_cross)))+
  set_rescor(FALSE),
  data = data_wider,
  data2 = list(Amat_cross = Amat_cross),
  chains = 2, cores = 2, iter = 1000
)
save(brms_biv_offspring_FFD, file = "output/models/brms_biv_offspring_FFD.rda")
```

### RUN: Separate for H and UH: Run more iterations!

#### RUNNING: Crosses

Warning message:
There were 342 divergent transitions after warmup. Increasing adapt_delta above 0.8 may help. See http://mc-stan.org/misc/warnings.html#divergent-transitions-after-warmup 

```{r eval=FALSE, include=FALSE}
brms_offpsring_FFD_full.h <- brm(
  FFD ~ 1 + (scale(temp_mother)+scale(temp_father))+
    (1 | gr(unique_id, cov = Amat_cross))+ (1 | block)+ (1|mother), 
  data = subset(data_exp_blocks,treat=="heatmat"),
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 4, cores = 16, iter = 5000, warmup = 1000, thin = 5
)
save(brms_offpsring_FFD_full.h, file = "output/models/brms_offpsring_FFD_full.h.rda")
```

```{r eval=FALSE, include=FALSE}
brms_offpsring_FFD_full.uh <- brm(
  FFD ~ 1 + (temp_mother+temp_father)+
    (1 | gr(unique_id, cov = Amat_cross))+ (1 | block) + (1|mother), 
  data = subset(data_exp_blocks,treat=="unheated"),
  data2 = list(Amat_cross = Amat_cross),
  family = gaussian(),
  chains = 2, cores = 1, iter = 1000
)
save(brms_offpsring_FFD_full.uh, file = "output/models/brms_offpsring_FFD_full.uh.rda")
```

```{r}
load("output/models/brms_offpsring_FFD_full.h.rda")
load("output/models/brms_offpsring_FFD_full.uh.rda")
```

Checks, etc.

Heritability, including fixed effects of origin temperature:

```{r}
v_fixed_offpsring_FFD_full.h<-rowVars(
  posterior_linpred(brms_offpsring_FFD_full.h,re_formula = NA))
v_animal_offpsring_FFD_full.h<-(VarCorr(brms_offpsring_FFD_full.h,
                                        summary=FALSE)$unique_id$sd)^2
v_block_offpsring_FFD_full.h<-(VarCorr(brms_offpsring_FFD_full.h,
                                     summary=FALSE)$block$sd)^2
v_mother_offpsring_FFD_full.h<-(VarCorr(brms_offpsring_FFD_full.h,
                                      summary=FALSE)$mother$sd)^2
v_r_offpsring_FFD_full.h<-(VarCorr(brms_offpsring_FFD_full.h,
                                   summary=FALSE)$residual$sd)^2
h2_offpsring_FFD_full.h<-as.mcmc(v_animal_offpsring_FFD_full.h/
                                  (v_fixed_offpsring_FFD_full.h+
                                     v_animal_offpsring_FFD_full.h+
                                     v_block_offpsring_FFD_full.h+
                                     v_mother_offpsring_FFD_full.h+
                                     v_r_offpsring_FFD_full.h))
m2_offpsring_FFD_full.h<-as.mcmc(v_mother_offpsring_FFD_full.h/
                                  (v_fixed_offpsring_FFD_full.h+
                                     v_animal_offpsring_FFD_full.h+
                                     v_block_offpsring_FFD_full.h+
                                     v_mother_offpsring_FFD_full.h+
                                     v_r_offpsring_FFD_full.h))
summary(h2_offpsring_FFD_full.h)
summary(m2_offpsring_FFD_full.h)
plot(h2_offpsring_FFD_full.h)
plot(m2_offpsring_FFD_full.h)
```

```{r}
v_fixed_offpsring_FFD_full.uh<-rowVars(
  posterior_linpred(brms_offpsring_FFD_full.uh,re_formula = NA))
v_animal_offpsring_FFD_full.uh<-(VarCorr(brms_offpsring_FFD_full.uh,
                                        summary=FALSE)$unique_id$sd)^2
v_block_offpsring_FFD_full.uh<-(VarCorr(brms_offpsring_FFD_full.uh,
                                     summary=FALSE)$block$sd)^2
v_mother_offpsring_FFD_full.uh<-(VarCorr(brms_offpsring_FFD_full.uh,
                                      summary=FALSE)$mother$sd)^2
v_r_offpsring_FFD_full.uh<-(VarCorr(brms_offpsring_FFD_full.uh,
                                   summary=FALSE)$residual$sd)^2
h2_offpsring_FFD_full.uh<-as.mcmc(v_animal_offpsring_FFD_full.uh/
                                  (v_fixed_offpsring_FFD_full.uh+
                                     v_animal_offpsring_FFD_full.uh+
                                     v_block_offpsring_FFD_full.uh+
                                     v_mother_offpsring_FFD_full.uh+
                                     v_r_offpsring_FFD_full.uh))
m2_offpsring_FFD_full.uh<-as.mcmc(v_mother_offpsring_FFD_full.uh/
                                  (v_fixed_offpsring_FFD_full.uh+
                                     v_animal_offpsring_FFD_full.uh+
                                     v_block_offpsring_FFD_full.uh+
                                     v_mother_offpsring_FFD_full.uh+
                                     v_r_offpsring_FFD_full.uh))
summary(h2_offpsring_FFD_full.uh)
summary(m2_offpsring_FFD_full.uh)
plot(h2_offpsring_FFD_full.uh)
plot(m2_offpsring_FFD_full.uh)
```

#### Variance components

##### UH treatment

```{r}
varcomps_offpsring_FFD_full.uh <- tibble(
  beta0 = posterior_samples(brms_offpsring_FFD_full.uh,pars="b_Intercept")[,1],
  VF= rowVars(posterior_linpred(brms_offpsring_FFD_full.uh, re_formula = NA)),
  VI = (VarCorr(brms_offpsring_FFD_full.uh,summary=F)$unique_id$sd[,"Intercept"])^2,
  VM = (VarCorr(brms_offpsring_FFD_full.uh,summary=F)$mother$sd[,"Intercept"])^2,
  VB= VarCorr(brms_offpsring_FFD_full.uh,summary=FALSE)$block$sd[,"Intercept"]^2,
  VR=(posterior_samples(brms_offpsring_FFD_full.uh,pars="sigma")[,1])^2
  ) %>% 
  mutate(VP = select(.,c(VF,VI,VM,VB,VR)) %>% rowSums()) %>% 
  mutate(.iteration = 1) %>% mutate(.iteration = cumsum(.iteration))
```

```{r}
varcomps_offpsring_FFD_full.uh %>% 
  pivot_longer(cols=c(VF:VR),names_to = "varcomp" , values_to = "variance") %>% 
  mutate(varcomp=fct_relevel(varcomp,c("VR","VI","VM","VB","VF"))) %>% 
  ggplot()+
  stat_halfeye(aes(y=varcomp,x=100*variance/VP),
                point_interval = mean_hdi,.width = 0.001,
                normalize="xy")+
  scale_x_continuous(name="Variance explained (%)",
                     lim=c(0,100),breaks=seq(from=0,to=100,by=10))+
  scale_y_discrete(name="Variance component",
                   labels=c(expression(italic(V[Residual])),
                            expression(italic(V[Individual])),
                            expression(italic(V[Mother])),
                            expression(italic(V[Block])),
                            expression(italic(V[Fixed(temp)])),
                            expression(italic(V[Fixed]))
                            ))+
  my_theme()+
  background_grid(colour.major = "grey95",colour.minor = "grey95")+
  ggtitle("Partitioning of total phenotypic variation for FFD (UH treatment)")+
  labs(caption = expression(paste("Total phenotypic variance =",italic(V[Fixed]),
                           " + ", italic(V[Block])," + ", italic(V[Mother]),
                           " + ", italic(V[Individual]),
                           " + ", italic(V[Residual])
                           )))
```

##### H treatment

```{r}
varcomps_offpsring_FFD_full.h <- tibble(
  beta0 = posterior_samples(brms_offpsring_FFD_full.h,pars="b_Intercept")[,1],
  VF= rowVars(posterior_linpred(brms_offpsring_FFD_full.h, re_formula = NA)),
  VI = (VarCorr(brms_offpsring_FFD_full.h,summary=F)$unique_id$sd[,"Intercept"])^2,
  VM = (VarCorr(brms_offpsring_FFD_full.h,summary=F)$mother$sd[,"Intercept"])^2,
  VB= VarCorr(brms_offpsring_FFD_full.h,summary=FALSE)$block$sd[,"Intercept"]^2,
  VR=(posterior_samples(brms_offpsring_FFD_full.h,pars="sigma")[,1])^2
  ) %>% 
  mutate(VP = select(.,c(VF,VI,VM,VB,VR)) %>% rowSums()) %>% 
  mutate(.iteration = 1) %>% mutate(.iteration = cumsum(.iteration))
```

```{r}
varcomps_offpsring_FFD_full.h %>% 
  pivot_longer(cols=c(VF:VR),names_to = "varcomp" , values_to = "variance") %>% 
  mutate(varcomp=fct_relevel(varcomp,c("VR","VI","VM","VB","VF"))) %>% 
  ggplot()+
  stat_halfeye(aes(y=varcomp,x=100*variance/VP),
                point_interval = mean_hdi,.width = 0.001,
                normalize="xy")+
  scale_x_continuous(name="Variance explained (%)",
                     lim=c(0,100),breaks=seq(from=0,to=100,by=10))+
  scale_y_discrete(name="Variance component",
                   labels=c(expression(italic(V[Residual])),
                            expression(italic(V[Individual])),
                            expression(italic(V[Mother])),
                            expression(italic(V[Block])),
                            expression(italic(V[Fixed(temp)])),
                            expression(italic(V[Fixed]))
                            ))+
  my_theme()+
  background_grid(colour.major = "grey95",colour.minor = "grey95")+
  ggtitle("Partitioning of total phenotypic variation for FFD (H treatment)")+
  labs(caption = expression(paste("Total phenotypic variance =",italic(V[Fixed]),
                           " + ", italic(V[Block])," + ", italic(V[Mother]),
                           " + ", italic(V[Individual]),
                           " + ", italic(V[Residual])
                           )))
```

# Session info

```{r}
sessionInfo()
```
