---
title: Small-scale genetic differentiation, in mean flowering time, but not in plasticity, along a geothermal heating gradient
subtitle: Data analyses
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 4
  word_document:
    toc: yes
    toc_depth: '4'
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(tibble.width = Inf)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Load the packages used

```{r load packages, include=FALSE}
library(tidyverse)
library(readxl)
library(ggeffects)
library(knitr)
library(lme4)
library(lmerTest)
library(gridExtra)
library(ggthemes)
library(sjPlot)
library(cowplot)
library(performance)
library(partR2)
library(ggh4x)
```

# Create custom ggplot themes

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

# Read data from csv files

```{r}
data_exp<-read_csv(file="data/clean/data_exp.csv")%>%
  mutate(across(c(unique_id, treat,sire,dam,dam_sire),as.factor),
         across(c(total_n_fl,FFD,LFD,MFD),as.integer),
         across(c(temp_dam,temp_sire,mean_sh_growth,mean_sh_h_date1,
                  mean_sh_h_date2),as.double))
```

# Data analyses

## Additive genetic variation in flowering time

### Models

```{r}
model1_FFD_uh<-lmer(FFD~1+(1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="unheated"))
model1_FFD_h<-lmer(FFD~1+(1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="heatmat"))
model1_LFD_uh<-lmer(LFD~1+(1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="unheated"))
model1_LFD_h<-lmer(LFD~1+(1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="heatmat"))
model1_MFD_uh<-lmer(MFD~1+(1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="unheated"))
model1_MFD_h<-lmer(MFD~1+(1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="heatmat"))
summary(model1_FFD_uh)
summary(model1_FFD_h)
summary(model1_LFD_uh)
summary(model1_LFD_h)
summary(model1_MFD_uh)
summary(model1_MFD_h)
```

### Proportions of variance

```{r}
Variance_FFD_uh<-as.data.frame(VarCorr(model1_FFD_uh))[,c(1,4)]
Variance_LFD_uh<-as.data.frame(VarCorr(model1_LFD_uh))[,c(1,4)]
Variance_MFD_uh<-as.data.frame(VarCorr(model1_MFD_uh))[,c(1,4)]
Variance_FFD_h<-as.data.frame(VarCorr(model1_FFD_h))[,c(1,4)]
Variance_LFD_h<-as.data.frame(VarCorr(model1_LFD_h))[,c(1,4)]
Variance_MFD_h<-as.data.frame(VarCorr(model1_MFD_h))[,c(1,4)]
# Intra-class correlations
PropVar_FFD_uh <- Variance_FFD_uh%>%mutate(propvar=vcov/sum(vcov))%>%
  mutate(variable="FFD",treat="unheated")
PropVar_LFD_uh <- Variance_LFD_uh%>%mutate(propvar=vcov/sum(vcov))%>%
  mutate(variable="LFD",treat="unheated")
PropVar_MFD_uh <- Variance_MFD_uh%>%mutate(propvar=vcov/sum(vcov))%>%
  mutate(variable="MFD",treat="unheated")
PropVar_FFD_h <- Variance_FFD_h%>%mutate(propvar=vcov/sum(vcov))%>%
  mutate(variable="FFD",treat="heated")
PropVar_LFD_h <- Variance_LFD_h%>%mutate(propvar=vcov/sum(vcov))%>%
  mutate(variable="LFD",treat="heated")
PropVar_MFD_h <- Variance_MFD_h%>%mutate(propvar=vcov/sum(vcov))%>%
  mutate(variable="MFD",treat="heated")
Props_var<-rbind(PropVar_FFD_uh,PropVar_LFD_uh,PropVar_MFD_uh,
                 PropVar_FFD_h,PropVar_LFD_h,PropVar_MFD_h)
```

```{r}
Props_var
```

### Heritability and maternal effects

```{r}
# h^2 (paternal effects)
her_FFD_uh<-as.numeric(4*subset(PropVar_FFD_uh,grp=="sire")[3])
her_FFD_h<-as.numeric(4*subset(PropVar_FFD_h,grp=="sire")[3])
her_LFD_uh<-as.numeric(4*subset(PropVar_LFD_uh,grp=="sire")[3])
her_LFD_h<-as.numeric(4*subset(PropVar_LFD_h,grp=="sire")[3])
her_MFD_uh<-as.numeric(4*subset(PropVar_MFD_uh,grp=="sire")[3])
her_MFD_h<-as.numeric(4*subset(PropVar_MFD_h,grp=="sire")[3])
# Because the additive genetic variance, VA,
# is expected to be four times the among pollen‐donor variance 
# (Falconer & Mackay, 1996; Lynch & Walsh, 1998)
her<-data.frame(value=rbind(her_FFD_uh,her_FFD_h,her_LFD_uh,
                       her_LFD_h,her_MFD_uh,her_MFD_h))%>%
  rownames_to_column()%>%
  mutate(variable=c("FFD","FFD","LFD","LFD","MFD","MFD"),
         treat=c("unheated","heated","unheated","heated",
                 "unheated","heated"),
         effect="Heritability")
```

```{r}
# maternal effects

# Maternal - paternal
# ----------------------
# summed effects

# Because the pollen‐recipient variance component contains a combination of
# genetic and environmental effects, we subtracted the additive genetic 
# (pollen donor) component from the pollen-recipient variance component
# before dividing the resulting estimate by VP to estimate 
# m2 (m2 = (Vpollen recipient − Vpollen donor)/VP).

mat_FFD_uh<-(Variance_FFD_uh[2,2]-Variance_FFD_uh[3,2])/
  as.numeric(Variance_FFD_uh%>%summarise(sum(vcov)))
mat_FFD_h<-(Variance_FFD_h[2,2]-Variance_FFD_h[3,2])/
  as.numeric(Variance_FFD_h%>%summarise(sum(vcov)))
mat_LFD_uh<-(Variance_LFD_uh[2,2]-Variance_LFD_uh[3,2])/
  as.numeric(Variance_LFD_uh%>%summarise(sum(vcov)))
mat_LFD_h<-(Variance_LFD_h[2,2]-Variance_LFD_h[3,2])/
  as.numeric(Variance_LFD_h%>%summarise(sum(vcov)))
mat_MFD_uh<-(Variance_MFD_uh[2,2]-Variance_MFD_uh[3,2])/
  as.numeric(Variance_MFD_uh%>%summarise(sum(vcov)))
mat_MFD_h<-(Variance_MFD_h[2,2]-Variance_MFD_h[3,2])/
  as.numeric(Variance_MFD_h%>%summarise(sum(vcov)))
mat<-data.frame(value=rbind(mat_FFD_uh,mat_FFD_h,mat_LFD_uh,
                       mat_LFD_h,mat_MFD_uh,mat_MFD_h))%>%
  rownames_to_column()%>%
  mutate(variable=c("FFD","FFD","LFD","LFD","MFD","MFD"),
         treat=c("unheated","heated","unheated","heated",
                 "unheated","heated"),
         effect="Maternal effects")
```

```{r}
her_mat<-rbind(her,mat)
her_mat
```

### Appendix S8: LRTs for variance components

```{r}
ranova(model1_FFD_uh)
ranova(model1_FFD_h)
ranova(model1_LFD_uh)
ranova(model1_LFD_h)
ranova(model1_MFD_uh)
ranova(model1_MFD_h)
```

## Additive genetic variation in flowering time plasticity

### Models

```{r}
model2_FFD<-lmer(FFD~treat+
                    (1|sire)+(1|dam)+
                    (1|treat:sire)+
                    #(1|treat:dam)+ # Removed to avoid singular fit
                    (1|sire:dam),
                  data_exp)
model2_LFD<-lmer(LFD~treat+
                    (1|sire)+(1|dam)+
                    #(1|treat:sire)+ # Removed to avoid singular fit
                    (1|treat:dam)+
                    (1|sire:dam),
                  data_exp)
model2_MFD<-lmer(MFD~treat+
                    (1|sire)+(1|dam)+
                    (1|treat:sire)+
                    #(1|treat:dam)+ # Removed to aovid singular fit
                    (1|sire:dam),
                  data_exp)
summary(model2_FFD)
summary(model2_LFD)
summary(model2_MFD)
```

### Proportions of variance

```{r}
# extract variance components
Variance_FFD_treat <- as.data.frame(VarCorr(model2_FFD))[,c(1,4)]  
Variance_LFD_treat <- as.data.frame(VarCorr(model2_LFD))[,c(1,4)]  
Variance_MFD_treat <- as.data.frame(VarCorr(model2_MFD))[,c(1,4)]  
# Intra-class correlation
PropVar_FFD_treat <- Variance_FFD_treat%>%
  mutate(propvar=vcov/sum(vcov))%>%mutate(variable="FFD")
PropVar_LFD_treat <- Variance_LFD_treat%>%
  mutate(propvar=vcov/sum(vcov))%>%mutate(variable="LFD")
PropVar_MFD_treat <- Variance_MFD_treat%>%
  mutate(propvar=vcov/sum(vcov))%>%mutate(variable="MFD")
# Proportional variance
Props_var_treat<-rbind(PropVar_FFD_treat,PropVar_LFD_treat,
                       PropVar_MFD_treat)
```

```{r}
Props_var_treat
```

### Heritability and maternal effects

```{r}
# h^2 (paternal effects)
her_FFD_treat<-4*subset(PropVar_FFD_treat,grp=="treat:sire")[3]
her_LFD_treat<-0
her_MFD_treat<-4*subset(PropVar_MFD_treat,grp=="treat:sire")[3]
# Because the additive genetic variance, VA,
# is expected to be four times the among pollen‐donor variance 
# (Falconer & Mackay, 1996; Lynch & Walsh, 1998)
her_treat<-data.frame(value=rbind(her_FFD_treat,
                                  her_LFD_treat,her_MFD_treat))%>%
  mutate(variable=c("FFD","LFD","MFD"),
         effect="Heritability")%>%
  rename(value=propvar)
```

```{r}
# maternal effects

# Maternal - paternal
# ----------------------
# summed effects

# Because the pollen‐recipient variance component contains a combination of
# genetic and environmental effects, we subtracted the additive genetic 
# (pollen donor) component from the pollen-recipient variance component
# before dividing the resulting estimate by VP to estimate 
# m2 (m2 = (Vpollen recipient − Vpollen donor)/VP).

mat_FFD_treat<-(0-
                  subset(Variance_FFD_treat,grp=="treat:sire")[2])/
  Variance_FFD_treat%>%summarise(sum(vcov))
mat_LFD_treat<-(subset(Variance_LFD_treat,grp=="treat:dam")[2]-
                  0)/
  Variance_LFD_treat%>%summarise(sum(vcov))
mat_MFD_treat<-(0-
                  subset(Variance_MFD_treat,grp=="treat:sire")[2])/
  Variance_MFD_treat%>%summarise(sum(vcov))
mat_treat<-data.frame(value=rbind(mat_FFD_treat,
                                  mat_LFD_treat,mat_MFD_treat))%>%
  mutate(variable=c("FFD","LFD","MFD"),
         effect="Maternal effects")%>%
  rename(value=vcov)
```

```{r}
her_mat_treat<-rbind(her_treat,mat_treat)
her_mat_treat
```

### Appendix S9: LRTs for variance components

```{r}
ranova(model2_FFD)
ranova(model2_LFD)
ranova(model2_MFD)
```

## Figure 1

```{r}
Props_var<-Props_var%>%
  mutate(variable=factor(variable,levels=c("FFD","MFD","LFD")),
         treat=factor(treat,levels=c("unheated","heated")),
         grp=factor(grp,levels=c("sire","dam","sire:dam","Residual")),
         sig=c("n","y","y","y",
               "n","y","n","Y",
               "n","y","y","y",
               "y","y","n","y",
               "n","y","n","y",
               "n","y","n","y"))%>% # Based on LRTs
  mutate(treat=fct_recode(treat,"control"="unheated"))
Props_var_treat<-Props_var_treat%>%
  mutate(variable=factor(variable,levels=c("FFD","MFD","LFD")),
         grp=factor(grp,levels=c("sire","dam","sire:dam",
                                 "treat:sire","treat:dam","Residual")),
         sig=c("n","y","y","n","y",
               "n","n","y","n","y",
               "n","y","y","n","y")) # Based on LRTs
```

```{r}
legend_fig1<-get_legend(
  ggplot(Props_var_treat%>%filter(grp!="Residual"),
         aes(x=variable,y=propvar,fill=grp,shape=sig))+
    geom_col(color="black",size=0.3)+
    geom_point(aes(x=variable,y=propvar,group=grp),
               position=position_stack(vjust=.5),show.legend=F,size=2)+
    my_theme_legend()+
    xlab(NULL)+ylab("Proportion of variance")+
    theme(legend.position="top",legend.title=element_blank())+
    scale_fill_discrete(labels=c("sire","dam","sire:dam",
                                 "treat:sire","treat:dam","residual"),
                        type=c("#00BFC4","#F8766D","#00BA38","#619CFF",
                               "#F564E3","grey"))+
    scale_shape_manual(values=c(NA,8))+ggtitle("B)")+
    theme(axis.text.y=element_blank(),axis.title.y=element_blank()))
fig1a<-ggplot(Props_var%>%filter(grp!="Residual"),
              aes(x=variable,y=propvar,fill=grp,shape=sig))+
  geom_col(color="black",size=0.3)+
  geom_point(aes(x=variable,y=propvar,group=grp),
             position=position_stack(vjust=.5),show.legend=F,size=2)+
  facet_wrap(~treat)+
  my_theme()+xlab(NULL)+ylab("Proportion of variance")+
  scale_fill_discrete(labels=c("sire","dam","sire:dam","residual"),
                      type=c("#00BFC4","#F8766D","#00BA38","grey"))+
  guides(fill=guide_legend(nrow=2))+
  theme(plot.title = element_text(hjust =-0.05,vjust=0))+
  scale_shape_manual(values=c(NA,8))+ggtitle("A)")
fig1b<-ggplot(Props_var_treat%>%filter(grp!="Residual"),
              aes(x=variable,y=propvar,fill=grp,shape=sig))+
  geom_col(color="black",size=0.3)+
  geom_point(aes(x=variable,y=propvar,group=grp),
             position=position_stack(vjust=.5),show.legend=F,size=2)+
  my_theme()+
  xlab(NULL)+ylab("Proportion of variance")+
  scale_fill_discrete(labels=c("sire","dam","sire:dam",
                               "treat:sire","treat:dam","residual"),
                      type=c("#00BFC4","#F8766D","#00BA38","#619CFF",
                             "#F564E3","grey"))+
  theme(plot.title = element_text(hjust =-0.05,vjust=4))+
  scale_shape_manual(values=c(NA,8))+ggtitle("B)")+
  theme(axis.title.y=element_blank())
fig1_without_legend<-cowplot::plot_grid(fig1a,fig1b,
                                        ncol=2,align="hv",axis="t",
                                        rel_widths = c(1.8, 1))
fig1<-grid.arrange(legend_fig1,fig1_without_legend,ncol=1,
                   heights=c(0.05,0.95))
ggsave(file="output/figures/fig1.tiff", plot=fig1,width=10,height=5,dpi=300,
       compression="lzw")
```

## Table 1

```{r}
her_mat<-her_mat%>%
  mutate(variable=factor(variable,levels=c("FFD","MFD","LFD")),
         treat=factor(treat,levels=c("unheated","heated")))%>%
  dplyr::select(-rowname)%>%
  pivot_wider(names_from="effect",values_from="value")
her_mat
her_mat_treat<-her_mat_treat%>%
  pivot_wider(names_from="effect",values_from="value")
her_mat_treat
```

## Figure 2

Main effect of temperature treatment significant for FFD, LFD and MFD.

```{r}
data_fig2<-rbind(
  data.frame(ggpredict(model2_FFD,terms="treat"))%>%mutate(variable="FFD"),
  data.frame(ggpredict(model2_MFD,terms="treat"))%>%mutate(variable="MFD"),
  data.frame(ggpredict(model2_LFD,terms="treat"))%>%mutate(variable="LFD"),
  data.frame(ggpredict(model2_FFD,terms=c("treat","sire"),type="random"))%>%
    mutate(variable="FFD"),
  data.frame(ggpredict(model2_MFD,terms=c("treat","sire"),type="random"))%>%
    mutate(variable="MFD"))%>%
  mutate(variable=factor(variable,levels=c("FFD","MFD","LFD")),
         x=factor(x,levels=c("unheated","heatmat"),
                  labels=c("unheated","heated")),
         type=ifelse(group==1,"mean","individual"))
```

```{r}
fig2<-ggplot()+
  # Mean responses
  geom_point(data=data_fig2%>%filter(group==1)%>%
               mutate(x=factor(x,levels=c("unheated","heated")))%>%
               mutate(x=fct_recode(x,"control"="unheated")),
             aes(x=x,y=predicted,color=variable),size=2)+
  geom_errorbar(data=data_fig2%>%filter(group==1)%>%
                  mutate(x=factor(x,levels=c("unheated","heated")))%>%
                  mutate(x=fct_recode(x,"control"="unheated")),
                aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high,
                    color=variable),
                width=0.05,size=0.5)+
  geom_line(data=data_fig2%>%filter(group==1)%>%
              mutate(x=factor(x,levels=c("unheated","heated")))%>%
              mutate(x=fct_recode(x,"control"="unheated")),
            aes(x=x,y=predicted,group=variable,color=variable),size=0.5)+
  # Response of each sire for FFD
  geom_line(data=data_fig2%>%filter(group!=1)%>%
              mutate(x=factor(x,levels=c("unheated","heated")))%>%
              mutate(x=fct_recode(x,"control"="unheated"))%>%
              filter(variable=="FFD"),
            aes(x=x,y=predicted,group=group),
            linewidth=0.2,color="#F8766D",alpha=0.5)+
  # Response of each sire for MFD
  geom_line(data=data_fig2%>%filter(group!=1)%>%
              mutate(x=factor(x,levels=c("unheated","heated")))%>%
              mutate(x=fct_recode(x,"control"="unheated"))%>%
              filter(variable=="MFD"),
            aes(x=x,y=predicted,group=group),
            linewidth=0.2,color="#00BA38",alpha=0.5)+
  my_theme_legend()+theme(legend.title=element_blank())+
  theme(plot.title = element_text(hjust =-0.2,vjust=-1.5))+
  theme(legend.position = c(0.84, 0.88))+
  theme(legend.background = element_rect(fill = "transparent"))+
  xlab(NULL)+ylab("Flowering date")+scale_x_discrete(expand=c(0.1,0.1))
ggsave(file="output/figures/fig2.tiff", plot=fig2,width=5,height=4,dpi=300,
       compression="lzw")
```

## Hypothesis 1

Temperature-dependent selection has resulted in small-scale genetic differentiation, and additive genetic variation in phenology is correlated with soil temperature in a counter-gradient fashion. At a given temperature, individuals with parents from warmer soils are predicted to start vegetative growth and flower later than plants with parents from colder soils.

### Models temperatures sire and dam

```{r}
model3_FFD_uh<-lmer(FFD~temp_sire+temp_dam+
                      (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="unheated"&!is.na(FFD)))
model3_FFD_h<-lmer(FFD~temp_sire+temp_dam+
                     (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="heatmat"&!is.na(FFD)))
model3_LFD_uh<-lmer(LFD~temp_sire+temp_dam+
                      (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="unheated"&!is.na(LFD)))
model3_LFD_h<-lmer(LFD~temp_sire+temp_dam+
                     (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="heatmat"&!is.na(LFD)))
model3_MFD_uh<-lmer(MFD~temp_sire+temp_dam+
                      (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="unheated"&!is.na(MFD)))
model3_MFD_h<-lmer(MFD~temp_sire+temp_dam+
                     #(1|sire) # Removed to avoid singular fit
                     (1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="heatmat"&!is.na(MFD)))
summary(model3_FFD_uh)
summary(model3_FFD_h)
summary(model3_LFD_uh)
summary(model3_LFD_h)
summary(model3_MFD_uh)
summary(model3_MFD_h)
```

#### Prediction results

```{r}
ggpredict(model3_MFD_h,terms=c("temp_sire[5]"))
ggpredict(model3_MFD_h,terms=c("temp_sire[45.5]"))
```

#### Table 2

```{r}
tab_model(model3_FFD_uh,model3_MFD_uh,model3_LFD_uh,
          transform=NULL,show.intercept=F,show.ci=F,show.se=T,show.stat=T,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("FFD","MFD","LFD"),
          file="output/tables/Table2_uh.html")
tab_model(model3_FFD_h,model3_MFD_h,model3_LFD_h,
          transform=NULL,show.intercept=F,show.ci=F,show.se=T,show.stat=T,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("FFD","MFD","LFD"),
          file="output/tables/Table2_h.html")
```

#### Proportions of variance

Proportion of variance explained by temperature of origin of sires and dams = proportion of variance explained by fixed effects: marginal R2.

```{r}
PropVar_FFD_uh_temp<-as.numeric(r2_nakagawa(model3_FFD_uh)$R2_marginal)
PropVar_FFD_h_temp<-as.numeric(r2_nakagawa(model3_FFD_h)$R2_marginal)
PropVar_LFD_uh_temp<-as.numeric(r2_nakagawa(model3_LFD_uh)$R2_marginal)
PropVar_LFD_h_temp<-as.numeric(r2_nakagawa(model3_LFD_h)$R2_marginal)
PropVar_MFD_uh_temp<-as.numeric(r2_nakagawa(model3_MFD_uh)$R2_marginal)
PropVar_MFD_h_temp<-as.numeric(r2_nakagawa(model3_MFD_h)$R2_marginal)
Props_var_temp<-data.frame(
  values=c(PropVar_FFD_uh_temp,PropVar_FFD_h_temp,
           PropVar_LFD_uh_temp,PropVar_LFD_h_temp,
           PropVar_MFD_uh_temp,PropVar_MFD_h_temp),
  variable=c("FFD","FFD","LFD","LFD","MFD","MFD"),
  treat=c("unheated","heated","unheated","heated",
                 "unheated","heated"))
```

```{r}
Props_var_temp
```

#### Temperature dam and sire: partR2

```{r}
partR2_model3_FFD_uh<-partR2(model3_FFD_uh,
                             partvars=c("temp_sire","temp_dam"),
                             R2_type="marginal",nboot=1000)
partR2_model3_FFD_h<-partR2(model3_FFD_h,
                             partvars=c("temp_sire","temp_dam"),
                             R2_type="marginal",nboot=1000)
partR2_model3_LFD_uh<-partR2(model3_LFD_uh,
                             partvars=c("temp_sire","temp_dam"),
                             R2_type="marginal",nboot=1000)
partR2_model3_LFD_h<-partR2(model3_LFD_h,
                             partvars=c("temp_sire","temp_dam"),
                             R2_type="marginal",nboot=1000)
partR2_model3_MFD_uh<-partR2(model3_MFD_uh,
                             partvars=c("temp_sire","temp_dam"),
                             R2_type="marginal",nboot=1000)
partR2_model3_MFD_h<-partR2(model3_MFD_h,
                             partvars=c("temp_sire","temp_dam"),
                             R2_type="marginal",nboot=1000)
```

```{r}
Props_var_temp_mf<-
  rbind(
    partR2_model3_FFD_uh$R2%>%mutate(variable="FFD",treat="unheated"),
    partR2_model3_FFD_h$R2%>%mutate(variable="FFD",treat="heated"),
    partR2_model3_LFD_uh$R2%>%mutate(variable="LFD",treat="unheated"),
    partR2_model3_LFD_h$R2%>%mutate(variable="LFD",treat="heated"),
    partR2_model3_MFD_uh$R2%>%mutate(variable="MFD",treat="unheated"),
    partR2_model3_MFD_h$R2%>%mutate(variable="MFD",treat="heated")
    )
```

```{r}
Props_var_temp_mf
```

#### Appendix S10

```{r}
AppS10<-ggplot()+
  geom_bar(data=Props_var_temp_mf%>%
             filter(term!="temp_sire+temp_dam"&term!="Full")%>%
             mutate(term=factor(term,levels=c("temp_sire","temp_dam"),
                                labels=c("Sire origin temperature",
                                           "Dam origin temperature")),
                    variable=factor(variable,levels=c("veg","FFD","MFD","LFD")),
                    sig=c("n","n","n","n",
                          "n","n","n","n","n","n","y","n"))%>%
               mutate(treat=fct_recode(treat,"control"="unheated")),
           aes(x=variable,y=estimate,fill=term),
           stat="identity",color="black",size=0.2)+
  facet_wrap(~factor(treat,levels=c("control","heated")))+
  geom_point(data=Props_var_temp_mf%>%
               filter(term!="temp_sire+temp_dam"&term!="Full")%>%
               mutate(term=factor(term,levels=c("temp_sire","temp_dam"),
                                  labels=c("Sire origin temperature",
                                           "Dam origin temperature")),
                      variable=factor(variable,
                                      levels=c("veg","FFD","MFD","LFD")),
                      sig=c("n","n","n","n",
                            "n","n","n","n","n","n","y","n"))%>%
               mutate(treat=fct_recode(treat,"control"="unheated")),
             aes(x=variable,y=estimate,group=term,shape=sig),
             position=position_stack(vjust=.5),show.legend=F,size=2)+
  my_theme_legend()+scale_fill_discrete(type=c("#00BFC4","#F8766D"))+
  scale_shape_manual(values=c(NA,8))+
  xlab(NULL)+ylab("Proportion of variance")+
  theme(legend.position="top",legend.title=element_blank())
AppS10
ggsave(file="output/figures/AppS10.tiff", plot=AppS10,width=5,height=5,dpi=300,
       compression="lzw")
```

#### Appendix S11

```{r}
AppS11<-ggplot(data=Props_var_temp_mf%>%
             filter(term!="temp_sire+temp_dam"&term!="Full")%>%
             mutate(term=factor(term,levels=c("temp_sire","temp_dam"),
                                labels=c("Sire origin temperature",
                                           "Dam origin temperature")),
                    variable=factor(variable,levels=c("FFD","MFD","LFD")),
                    sig=c("n","n","n","n",
                          "n","n","n","n","n","n","y","n"))%>%
               mutate(treat=fct_recode(treat,"control"="unheated")),
           aes(x=variable,y=estimate,ymin=CI_lower,ymax=CI_upper,color=term))+
  geom_point(size=2,position=position_dodge(0.4),width=0.2)+
  geom_errorbar(position=position_dodge(0.4),width=0.2)+
  facet_wrap(~factor(treat,levels=c("control","heated")))+
  my_theme_legend()+
  scale_color_discrete(type=c("#00BFC4","#F8766D"))+
  xlab(NULL)+ylab("Proportion of variance")+
  theme(legend.position="top",legend.title=element_blank())
AppS11
ggsave(file="output/figures/AppS11.tiff", plot=AppS11,width=7,height=8,dpi=300,
       compression="lzw")
```

### Models mid-parental values

```{r}
data_exp<-data_exp%>%
  mutate(temp_midP=(temp_sire+temp_dam)/2)
```

```{r}
model3_FFD_uh_midP<-lmer(FFD~temp_midP+
                      (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="unheated"&!is.na(FFD)))
model3_FFD_h_midP<-lmer(FFD~temp_midP+
                     (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="heatmat"&!is.na(FFD)))
model3_LFD_uh_midP<-lmer(LFD~temp_midP+
                      (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="unheated"&!is.na(LFD)))
model3_LFD_h_midP<-lmer(LFD~temp_midP+
                     (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="heatmat"&!is.na(LFD)))
model3_MFD_uh_midP<-lmer(MFD~temp_midP+
                      (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="unheated"&!is.na(MFD)))
model3_MFD_h_midP<-lmer(MFD~temp_midP+
                     (1|sire)+(1|dam)+(1|sire:dam),
                 subset(data_exp,treat=="heatmat"&!is.na(MFD)))
summary(model3_FFD_uh_midP)
summary(model3_FFD_h_midP)
summary(model3_LFD_uh_midP)
summary(model3_LFD_h_midP)
summary(model3_MFD_uh_midP)
summary(model3_MFD_h_midP)
```

#### Appendix S6

```{r}
tab_model(model3_FFD_uh_midP,model3_MFD_uh_midP,
          model3_LFD_uh_midP,
          transform=NULL,show.intercept=F,show.ci=F,show.se=T,show.stat=T,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("FFD","MFD","LFD"),
          file="output/tables/AppS6_uh.html")
tab_model(model3_FFD_h_midP,model3_MFD_h_midP,
          model3_LFD_h_midP,
          transform=NULL,show.intercept=F,show.ci=F,show.se=T,show.stat=T,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("FFD","MFD","LFD"),
          file="output/tables/AppS6_h.html")
```

## Hypothesis 2

Genetic differentiation related to soil temperature involves also plasticity of development rate. Differences in phenology between high and low temperatures are predicted to be smaller for plants with parents from warm soils than for plants from colder soils.

### Models temperatures sire and dam

```{r}
model4_FFD<-lmer(FFD~treat*(temp_sire+temp_dam)+
                    (1|sire)+(1|dam)+
                    (1|sire:dam),
                  data_exp)
model4_LFD<-lmer(LFD~treat*(temp_sire+temp_dam)+
                    (1|sire)+(1|dam)+
                    (1|sire:dam),
                  data_exp)
model4_MFD<-lmer(MFD~treat*(temp_sire+temp_dam)+
                    (1|sire)+(1|dam)+
                    (1|sire:dam),
                  subset(data_exp,!is.na(MFD)))
summary(model4_FFD)
summary(model4_LFD)
summary(model4_MFD)
```

#### Prediction results

```{r}
ggpredict(model4_MFD,terms=c("temp_sire[5]"))
ggpredict(model4_MFD,terms=c("temp_sire[45.5]"))
```

#### Table 3

```{r}
tab_model(model4_FFD,model4_MFD,model4_LFD,
          transform=NULL,show.intercept=F,show.ci=F,show.se=T,show.stat=T,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("FFD","MFD","LFD"),
          file="output/tables/Table3.html")
```

#### Figure 3

```{r}
fig3<-ggplot()+
  geom_ribbon(data=data.frame(ggpredict(model3_MFD_h,
                                        terms=c("temp_sire"))),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="red",alpha=0.25)+
  geom_line(data=data.frame(ggpredict(model3_MFD_uh,
                                      terms=c("temp_sire"))),
            aes(x=x,y=predicted),color="blue",linetype="dashed")+
  geom_ribbon(data=data.frame(ggpredict(model3_MFD_uh,
                                        terms=c("temp_sire"))),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="blue",alpha=0.25)+
  geom_line(data=data.frame(ggpredict(model3_MFD_h,
                                      terms=c("temp_sire"))),
            aes(x=x,y=predicted),color="red")+
  geom_ribbon(data=data.frame(ggemmeans(model4_MFD,
                                        terms=c("temp_sire"))),
              aes(x=x,y=predicted,ymin=conf.low,ymax=conf.high),
              fill="darkgrey",alpha=0.25)+
  geom_line(data=data.frame(ggemmeans(model4_MFD,
                                      terms=c("temp_sire"))),
            aes(x=x,y=predicted),color="black",size=0.75)+
  geom_jitter(data=data_exp%>%
                mutate(treat=fct_recode(treat,"control"="unheated")),
              aes(x=temp_sire,y=MFD,color=treat),size=1,alpha=0.1)+
  scale_color_manual(values=c("blue","red"),limits=c("control","heatmat"),
                     labels=c("control","heated"))+
  my_theme_legend()+xlab("Sire origin temperature (ºC)")+
  ylab("Median flowering date")+
  theme(legend.position="top",legend.title=element_blank())+
  theme(legend.position = c(0.25, 0.95),legend.direction="horizontal")+
  theme(legend.background = element_rect(fill = "transparent"),
        legend.spacing.x = unit(0.1, 'cm'))
fig3
ggsave(file="output/figures/fig3.tiff", plot=fig3,width=4,height=4,dpi=300,
       compression="lzw")
```

### Models mid-parental values

```{r}
model4_FFD_midP<-lmer(FFD~treat*temp_midP+
                    (1|sire)+(1|dam)+
                    (1|sire:dam),
                  data_exp)
model4_LFD_midP<-lmer(LFD~treat*temp_midP+
                    (1|sire)+(1|dam)+
                    (1|sire:dam),
                  data_exp)
model4_MFD_midP<-lmer(MFD~treat*temp_midP+
                    (1|sire)+(1|dam)+
                    (1|sire:dam),
                  data_exp)
summary(model4_FFD_midP)
summary(model4_LFD_midP)
summary(model4_MFD_midP)
```

#### Appendix S7

```{r}
tab_model(model4_FFD_midP,model4_MFD_midP,model4_LFD_midP,
          transform=NULL,show.intercept=F,show.ci=F,show.se=T,show.stat=T,
          show.r2=T,show.icc=F,show.re.var=F,show.ngroups=F,digits=3,
          dv.labels=c("FFD","MFD","LFD"),
          file="output/tables/AppS7.html")
```

# Appendix S12: relations with flower number

```{r}
data_exp<-data_exp%>%mutate(total_n_fl_log=log(total_n_fl))
```

```{r}
model_FFD_fln<-lm(FFD~total_n_fl_log,data_exp%>%filter(total_n_fl>0))
model_MFD_fln<-lm(MFD~total_n_fl_log,data_exp%>%filter(total_n_fl>0))
model_LFD_fln<-lm(LFD~total_n_fl_log,data_exp%>%filter(total_n_fl>0))
```

```{r}
summary(model_FFD_fln)
summary(model_MFD_fln)
summary(model_LFD_fln)
```

```{r}
data_plot_fln<-rbind(
  data.frame(ggpredict(model_FFD_fln))%>%mutate(variable="FFD"),
  data.frame(ggpredict(model_MFD_fln))%>%mutate(variable="MFD"),
  data.frame(ggpredict(model_LFD_fln))%>%mutate(variable="LFD")
)%>%
  mutate(variable=factor(variable,levels = c("FFD", "MFD", "LFD")))
```

```{r}
r2_values <- data.frame(
  variable = c("FFD", "MFD", "LFD"),
  r2 = c(round(summary(model_FFD_fln)$r.squared,3),
         round(summary(model_MFD_fln)$r.squared,3),
         round(summary(model_LFD_fln)$r.squared,3))
)
```

```{r}
data_plot_fln <- data_plot_fln %>%
  left_join(r2_values, by = "variable")
```

```{r}
custom_labeller <- as_labeller(
  setNames(
    paste0(c("FFD", "MFD", "LFD"), "\nR² = ", c(0.362, 0.040, 0.107)), 
    c("FFD", "MFD", "LFD")
    )
  )
```


```{r}
ggplot()+
  geom_point(data=data_exp%>%select(FFD,LFD,MFD,total_n_fl_log)%>%
               pivot_longer(cols=c(FFD, MFD, LFD),names_to="variable",
                            values_to = "value")%>%
               mutate(variable=factor(variable,levels=c("FFD","MFD","LFD"))),
             aes(x=total_n_fl_log,y=value),
             alpha=0.25,size=1.5,show.legend=F)+
  facet_wrap(~variable, labeller = custom_labeller) + 
  geom_ribbon(data=data_plot_fln,
              aes(x=total_n_fl_log.x,y=total_n_fl_log.predicted,
                  ymin=total_n_fl_log.conf.low,ymax=total_n_fl_log.conf.high),
              alpha=0.25)+
  geom_line(data=data_plot_fln,
            aes(x=total_n_fl_log.x,y=total_n_fl_log.predicted))+
  my_theme()+theme(legend.title=element_blank())+
  xlab("Total number of flowers (log)")+ylab("Flowering date")
ggsave(file="output/figures/AppS12.tiff", width=12,height=5,dpi=300,
       compression="lzw")
```

# Appendix S4: Distances crossings

Read data from csv files:

```{r}
distances_crossings<-read_csv(file="data/clean/distances_crossings.csv")
distances_crossings
```

Histograms:

```{r}
new_new_appendix<-grid.arrange(
  ggplot(distances_crossings,aes(x=dist))+
    geom_histogram(color="black",fill="grey",bins=15)+
    my_theme()+xlab("Distance (m)")+ylab("Number of crossings")+ggtitle("A)"),
  ggplot(data_exp%>%mutate(temp_diff=abs(temp_dam-temp_sire))%>%
           distinct(dam_sire,temp_diff),
         aes(x=temp_diff))+geom_histogram(color="black",fill="grey",bins=15)+
    my_theme()+
    xlab("Absolute temperature difference (ºC)")+
    ylab("Number of crossings")+ggtitle("B)"),
  ncol=2)
ggsave(file="output/figures/new_new_appendix.tiff", plot=new_new_appendix,
       width=12,height=5,dpi=300,compression="lzw")
```

Ranges, means and SDs (for main text):

```{r}
range(distances_crossings$dist)
mean(distances_crossings$dist)
sd(distances_crossings$dist)
```
```{r}
range((data_exp%>%mutate(temp_diff=abs(temp_dam-temp_sire))%>%
         distinct(dam_sire,temp_diff))$temp_diff)
mean((data_exp%>%mutate(temp_diff=abs(temp_dam-temp_sire))%>%
         distinct(dam_sire,temp_diff))$temp_diff)
sd((data_exp%>%mutate(temp_diff=abs(temp_dam-temp_sire))%>%
         distinct(dam_sire,temp_diff))$temp_diff)
```

# Session info

```{r}
sessionInfo()
```
