---
title: Small-scale genetic differentiation, in mean flowering time, but not in plasticity, along a geothermal heating gradient
subtitle: Data analyses
author: "Alicia Valdés"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '4'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 4
  word_document:
    toc: yes
    toc_depth: '4'
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

```{r load packages, include=FALSE}
library(tidyverse)
library(sf)
library(ggmap)
library(raster)
library(here)
library(readxl)
library(ggthemes)
library(ggrepel)
library(viridis)
library(ggspatial)
```

```{r Define ggplot themes and palettes, include=FALSE}
my_theme <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(legend.position="none")+theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
my_theme_legend <- function(){
  theme_base()+theme(plot.background=element_rect(fill="white", colour=NA))+
  theme(text=element_text(family="serif"))+
  theme(plot.title = element_text(hjust =-0.06))
}
```

??read# Read shapefiles plants and plots

```{r}
plants_2017 <- st_read("data/shapefiles/plants_2017.shp")
plots <- st_read("data/shapefiles/plots.shp")
```

# Read data mothers and fathers

```{r}
data_mothers_fathers<-read_excel(
  "data/edited/Cerastium_greenhouse_spring_2022_editedAV.xlsx", 
  sheet="extracted_data")%>%
  dplyr::select(mother,father,temp_mother,temp_father,mother_originalID_iceland,
                father_originalID_iceland)%>%
  distinct() # Removes duplicate rows
```

## Change incorrect temperature

Temperature for mother 28 should be 45.5, not 25.5.

```{r}
data_mothers_fathers <- data_mothers_fathers %>%
  mutate(temp_mother = if_else(mother == 28, 45.5, temp_mother))
```

# Read data experiment

```{r}
data_exp<-read_csv(file="data/clean/data_exp.csv")%>%
  dplyr::select(dam,sire)%>%distinct()
```

# Get list of parent plants used in crossings

Get unique mothers and unique fathers:

```{r}
unique_mothers<-unique(data_exp$dam)
unique_fathers<-unique(data_exp$sire)
```

```{r}
# Reshape the data for mothers
mothers <- data_mothers_fathers %>%
  dplyr::select(mother, temp_mother, mother_originalID_iceland) %>%
  filter(mother %in% unique_mothers)%>%
  rename(newID = mother, temperature = temp_mother, 
         originalID = mother_originalID_iceland) %>%
  mutate(use_in_crossing = "mother")
# Reshape the data for fathers
fathers <- data_mothers_fathers %>%
  dplyr::select(father, temp_father, father_originalID_iceland) %>%
  filter(father %in% unique_fathers)%>%
  rename(newID = father, temperature = temp_father, 
         originalID = father_originalID_iceland) %>%
  mutate(use_in_crossing = "father")
# Combine the reshaped data
data_parents<-bind_rows(mothers,fathers)%>%
  distinct()%>% # Removes duplicate rows
  # split  originalID into plot and pl_id
  mutate(plot = str_extract(originalID, "^\\w+"), 
         # "^\\w+" matches the initial word part of originalID.
         pl_id = as.integer(str_extract(originalID, "(?<=\\s)\\d+")))%>%
         # "(?<=\\s)\\d+" matches the digits that follow a space in originalID
  # Creamos una columna que combine newID con info on if mother or father
  mutate(id_crossing=paste0(ifelse(use_in_crossing=="mother","D","S"),newID))%>%
  # Remove rows with NAs
  filter(!is.na(newID))%>%
  mutate(plot_pl_id = paste0(plot, "_", pl_id))
data_parents
```

Plants with several records:

```{r}
data_parents%>%group_by(plot_pl_id)%>%summarise(count=n())%>%filter(count>1)
```

# Select in shapefile plants that are on list of parent plants

```{r}
# Filter the 2017 shapefile to include only those rows in data_parents
plants_2017_parents <- plants_2017 %>%
  mutate(pl_id=as.integer(str_replace(pl_id,"H","")))%>%
  mutate(plot_pl_id = paste0(plot, "_", pl_id))%>%
  filter(plot_pl_id %in% data_parents$plot_pl_id)
```

Parent plants that are not in the 2017 shapefile: None!

```{r}
setdiff(data_parents$plot_pl_id, plants_2017_parents$plot_pl_id)
```
# Plot the shapefile of parent plants

```{r}
ggplot(data=plants_2017_parents)+geom_sf()+my_theme()
```

# Create shapefile with plot centroids and plot

```{r}
plot_centroids <- plants_2017_parents %>%
  group_by(plot) %>%
  summarize(geometry = st_union(geometry))%>%
  st_centroid() %>%
  st_as_sf()
```

```{r}
ggplot(data=plot_centroids)+geom_sf()+my_theme()+
  geom_text(aes(x = st_coordinates(geometry)[, 1], 
                y = st_coordinates(geometry)[, 2], 
                label = plot), size = 3, nudge_y = 0.0002)+
  xlab("X")+ylab("Y")
```

# Map of plots with satellite image

In QGIS!

# Maps of each plot

Add temperature data and newIDs:

```{r}
plants_2017_parents<-plants_2017_parents%>%
  left_join(data_parents%>%dplyr::select(-originalID))
```

Create three columns for id_crossing, as some plants are used two or three times:

```{r}
plants_2017_parents_merged<-plants_2017_parents%>%
  dplyr::select(plot,pl_id,plot_pl_id,temperature,id_crossing)%>%
  group_by(plot,pl_id) %>%
  mutate(id_crossing_1 = id_crossing[1],
         id_crossing_2 = ifelse(n() > 1, id_crossing[2], NA_character_),
         id_crossing_3 = ifelse(n() > 2, id_crossing[3], NA_character_))%>%
  dplyr::select(-id_crossing)%>%
  slice(1)
```

Create labels with id_crossing and temperature

```{r}
plants_2017_parents_merged<-plants_2017_parents_merged%>%
  mutate(ids_crossing=ifelse(
    !is.na(id_crossing_3),
    paste(id_crossing_1,id_crossing_2,id_crossing_3,sep="/"),
    ifelse(!is.na(id_crossing_2),
           paste(id_crossing_1,id_crossing_2,sep="/"),
           id_crossing_1)),
    label=paste(ids_crossing,paste(temperature,"ºC"),sep="\n"))
```


#